---
title: "Coxph Report nAb"
author: "CoVPN 3008"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Change the directory as needed
setwd('Github/antibody_manuscript/Cox_HR_nAb/')
set.seed(1)
setwd("C:/Users/bzhang3/Dropbox/3008 correlates/manuscript/antibody_manuscript/NAb_curves/")
```


```{r load pckages, echo=FALSE, warning=FALSE, message=FALSE}
library(CaseCohortCoxSurvival)
library(survival)
library(dplyr)
library(survey)
library(ggplot2)
library(GGally)
library(vaccine)
library(wCorr)
library(haven)
library(Hmisc)
source('CVE_functions.R')
```


# Data preprocessing

```{r data wrangling, echo=FALSE}
# Read in the processed dataset
dt = read.csv('./COVID_realdata_ics_bama_nab_bcp20241210_processed_with_riskscore.csv')

# Additional data processing
# Define log10-transformed marker variables
# If a participant is not in the phase-2 sample, then set the 
# participant's immune marker to NA
dt_serum_PP = dt %>%
  filter(ph1.m1.sera == 1) %>%
  mutate(hybrid = AGRP2N %in% c(2.1, 4.1) + 0) %>%
  mutate(is.subcohort = ifelse(IMMFL == 'Y', 1, 0)) %>%
  mutate(log10M1BA45ID50 = log10(M1BA45ID50),
         log10BBA45ID50 = log10(BBA45ID50),
         frBA45ID50 = log10M1BA45ID50 - log10BBA45ID50,
         log10M1IGGSpikeIndex = log10(M1IGGMSDA16AU), #A16 = IgG S Index
         log10BIGGSpikeIndex = log10(BIGGMSDA16AU), 
         frIGGSpikeIndex = log10M1IGGSpikeIndex - log10BIGGSpikeIndex,
         log10M1IGGSpikeBA45 = log10(M1IGGMSDA23AU), # A23 = IgG S BA.4/5
         log10BIGGSpikeBA45 = log10(BIGGMSDA23AU), 
         frIGGSpikeBA45 = log10M1IGGSpikeBA45 - log10BIGGSpikeBA45,
         log10M1IGGN = log10(M1IGGMSDA15AU),
         log10BIGGN = log10(BIGGMSDA15AU))

# Define a stratification variable W corresponding to
# 4 baseline strata + case
# Mutate ph2 to logical 
dt_serum_PP = dt_serum_PP %>%
  mutate(W = ifelse(EventIndPrimary1 == 1, 1, 
                    ifelse(AGRP2N == 1, 2, 
                           ifelse(AGRP2N == 2.1, 3,
                                  ifelse(AGRP2N == 3, 4, 5))))) %>%
  mutate(ph2.m1.sera_logical = ifelse(ph2.m1.sera == 1, TRUE, FALSE))

# Additional data processing
# Derive baseline variables to be adjusted for in the 
# antibody immune correlates analysis
dt_serum_PP = dt_serum_PP %>%
  mutate(TBFL = ifelse(TBFL == '', 'N', TBFL),
         TBFL = ifelse(TBFL == 'Y', 1, 0)) %>%
  mutate(AGEGR1 = ifelse(AGEGR1 == '>40', 1, 0),
         BMIB2GR1 = ifelse(BMIB2GR1 == '>25', 1, 0),
         CD4B2GR1 = ifelse(is.na(CD4B2GR1), 'NotKnown', 
                           ifelse(CD4B2GR1 == '<=500', 'low', 'high')),
         riskscore = ifelse(is.na(riskscore), median(riskscore, na.rm = T), riskscore),
         enrollmentPeriodEarly = ifelse(CalendarDateEnrollment <= 90, 1, 0),
         enrollmentPeriodMid = ifelse((CalendarDateEnrollment > 90 & CalendarDateEnrollment <= 180), 1, 0),
         regionRSA = ifelse(REGION == 'South Africa', 1, 0))
```

## Define serum markers
```{r setmarkerofinterest, echo=TRUE}
antigen_vec = c('BA45ID50', 'IGGSpikeIndex', 'IGGSpikeBA45', 'IGGN')
antigen_names_vec = c('nAb ID50 BA.4/5', 'IgG S Index',  
                      'IgG Spike BA.4/5', 'IgG N Index')
mk_vec_b = paste0('log10B', antigen_vec)
names_vec_b = c('M0 nAb-ID50 BA.4/5 (AU/ml)',
              'M0 IgG S Index (AU/ml)',
              'M0 IgG Spike BA.4/5 (AU/ml)',
              'M0 IgG N Index (AU/ml)')

mk_vec_p = paste0('log10M1', antigen_vec)
names_vec_p = c('Peak nAb-ID50 BA.4/5 (AU/ml)',
              'Peak IgG S Index (AU/ml)',
              'Peak IgG Spike BA.4/5 (AU/ml)',
              'Peak IgG N Index (AU/ml)')

mk_vec_fr = c('frBA45ID50', 'frIGGSpikeBA45')
names_vec_fr = c('Fold-rise nAb-ID50 BA.4/5 (AU/ml)',
              'Fold-rise IgG Spike BA.4/5 (AU/ml)')
```


## Define adjusted covariates
```{r adj covs}
adj_vars = c('regionRSA', 'HIVBL', 'riskscore', 
             'enrollmentPeriodEarly') 
n_var = 4
```


# Hybrid immunity group controlled risk and relative risk analysis


```{r cohort HIPP}
dt_serum_PP_HIPP = dt_serum_PP %>%
  filter(hybrid == 1)
```

## Plot the controlled VE curve for baseline and peak markers

```{r controlled VE HIPP, echo=FALSE, message=FALSE, warning=FALSE}
controlled_ve_mk_vec = c(mk_vec_b, mk_vec_p)
controlled_ve_names_vec = c(names_vec_b, names_vec_p)

a = length(controlled_ve_mk_vec)
for (i in 1:a){
  mk = controlled_ve_mk_vec[i]
  cat(mk, '\n')
  dat <- load_data(
  time = 'EventTimePrimary',
  event = 'EventIndPrimary1',
  vacc = 'hybrid',
  marker = mk,
  covariates = adj_vars,
  weights = "wt.m1.sera",
  ph2 = "ph2.m1.sera",
  data = dt_serum_PP)

b = vaccine::est_overall(dat, t_0 = 165)

overall_incidence = b$est[b$group == 'vaccine']  

ests_cox <- est_ce(dat = dat, type="Cox", t_0 = 165, return_p_value = TRUE, 
                   cve = TRUE, placebo_risk_method = 'Cox')
ests_np <- est_ce(dat = dat, type="NP", t_0 = 165, return_p_value = TRUE, 
                  cve = TRUE, placebo_risk_method = 'Cox')

df_np = data.frame(marker_level =  ests_np$cr$s,
                      controlled_risk =  ests_np$cr$est, 
                      ci_lower =  ests_np$cr$ci_lower, 
                      ci_upper =  ests_np$cr$ci_upper) 
write.csv(df_np, paste0( './source_data/Fig3_HIPP_',  mk, '.csv'))

mk_dist = dt_serum_PP_HIPP %>%
  filter(ph2.m1.sera == 1) %>%
  select(!!sym(mk), wt.m1.sera)

mk_dist_VIPP = dt_serum_PP %>%
  filter(hybrid == 0) %>%
  filter(ph2.m1.sera == 1) %>%
  select(!!sym(mk), wt.m1.sera)

# Weighted median of the marker level among VIPP
mk_VIPP_med = wtd.quantile(mk_dist_VIPP[,1], 
                           weights = mk_dist_VIPP[,2],
                           probs = c(0.5))

p = plot_cr2(ests_cox, ests_np, mk_dist, overall_incidence,
                     xlab_name = controlled_ve_names_vec[i],
                     title_name = 'Marker-Specific Risk',
                     t0 = 165, x_range = c(-0.5, 7.5),
                     x_breaks = seq(0, 7, 1),
                     x_labels =  c(expression(10^0), expression(10^1), expression(10^2),
                                   expression(10^3), expression(10^4), expression(10^5),
                                   expression(10^6), expression(10^7)))
p = p + geom_line(linewidth = 1.1) + 
  geom_ribbon(ggplot2::aes(ymin = ci_lower, ymax = ci_upper), 
              alpha = 0.2, linewidth = 1.1, linetype = "dotted") +
  theme_bw(base_size = 30) + 
  theme(legend.position = 'none') + ggtitle('')

saveRDS(p, paste0('./raw_data_for_figure/', controlled_ve_mk_vec[i], '_HIPP.RData'))

ggsave(paste0('./Figure/', controlled_ve_mk_vec[i], '_HIPP.pdf'),
       p, width = 12, height = 8, units = 'in')
}
```


## Plot the controlled relative risk curve for M0 markers: HIPP vs VIPP 

```{r controlled RR HIPP, echo=FALSE, message=FALSE, warning=FALSE}
controlled_ve_mk_vec = c(mk_vec_b)
controlled_ve_names_vec = c(names_vec_b)

a = length(controlled_ve_mk_vec)
for (i in c(1,3,4)){
  mk = controlled_ve_mk_vec[i]
  cat(mk, '\n')
  dat <- load_data(
  time = 'EventTimePrimary',
  event = 'EventIndPrimary1',
  vacc = 'hybrid',
  marker = mk,
  covariates = adj_vars,
  weights = "wt.m1.sera",
  ph2 = "ph2.m1.sera",
  data = dt_serum_PP)

b = vaccine::est_overall(dat, t_0 = 165)

overall_incidence = b$est[b$group == 'vaccine']  

ests_cox <- est_ce(dat = dat, type="Cox", t_0 = 165, return_p_value = TRUE, 
                   cve = TRUE, placebo_risk_method = 'Cox')
ests_np <- est_ce(dat = dat, type="NP", t_0 = 165, return_p_value = TRUE, 
                  cve = TRUE, placebo_risk_method = 'Cox')



ests_np2 = ests_np
ests_np2$cve$est = 1 - ests_np$cve$est
ests_np2$cve$ci_lower = 1 - ests_np$cve$ci_upper
ests_np2$cve$ci_upper = 1 - ests_np$cve$ci_lower

df_np = data.frame(marker_level =  ests_np2$cve$s,
                      RR =  ests_np2$cve$est, 
                      ci_lower =  ests_np2$cve$ci_lower, 
                      ci_upper =  ests_np2$cve$ci_upper) 
write.csv(df_np, paste0( './source_data/SuppFig18_',  mk, '.csv'))

mk_dist = dt_serum_PP_HIPP %>%
  filter(ph2.m1.sera == 1) %>%
  select(!!sym(mk), wt.m1.sera)

mk_dist_VIPP = dt_serum_PP %>%
  filter(hybrid == 0) %>%
  filter(ph2.m1.sera == 1) %>%
  select(!!sym(mk), wt.m1.sera)

# Weighted median of the marker level among VIPP
mk_VIPP_med = wtd.quantile(mk_dist_VIPP[,1], 
                           weights = mk_dist_VIPP[,2],
                           probs = c(0.5))

p = plot_cve2(ests_cox, ests_np2, mk_dist, mk_VIPP_med,
                     xlab_name = controlled_ve_names_vec[i],
                     title_name = 'Marker-Specific Relative Risk',
                     t0 = 165, x_range = c(-0.5, 7.5),
                     x_breaks = seq(0, 7, 1),
                     x_labels =  c(expression(10^0), expression(10^1), expression(10^2),
                                   expression(10^3), expression(10^4), expression(10^5),
                                   expression(10^6), expression(10^7)))

p = p + geom_line(linewidth = 1.1) + 
  geom_ribbon(ggplot2::aes(ymin = ci_lower, ymax = ci_upper), 
              alpha = 0.2, linewidth = 1.1, linetype = "dotted") +
  theme_bw(base_size = 30) + 
  theme(legend.position = 'none') + ggtitle('')

saveRDS(p, paste0('./raw_data_for_figure/', controlled_ve_mk_vec[i], '_RR_HIPP.RData'))

ggsave(paste0('./Figure/', controlled_ve_mk_vec[i], '_RR_HIPP.pdf'),
       p, width = 12, height = 8, units = 'in')
}
```

## Vaccine immunity group controlled risk and relative risk analysis

```{r cohort VIPP}
dt_serum_PP_VIPP = dt_serum_PP %>%
  filter(hybrid == 0) %>%
  mutate(vaccine = 1)
```

### Plot the controlled risk curve for peak markers

```{r controlled VE VIPP, echo=FALSE, message=FALSE, warning=FALSE}
controlled_ve_mk_vec = c(mk_vec_p)
controlled_ve_names_vec = c(names_vec_p)

a = length(controlled_ve_mk_vec)
for (i in 1:a){
  mk = controlled_ve_mk_vec[i]  
  cat(mk, '\n')
  dat <- load_data(
  time = 'EventTimePrimary',
  event = 'EventIndPrimary1',
  vacc = 'vaccine',
  marker = mk,
  covariates = adj_vars,
  weights = "wt.m1.sera",
  ph2 = "ph2.m1.sera",
  data = dt_serum_PP_VIPP)

b = vaccine::est_overall(dat, t_0 = 165, ve = FALSE)

overall_incidence = b$est[b$group == 'vaccine']  

ests_cox <- est_ce(dat = dat, type="Cox", t_0 = 165, return_p_value = TRUE, 
                   cve = FALSE)
ests_np <- est_ce(dat = dat, type="NP", t_0 = 165, return_p_value = TRUE, 
                  cve = FALSE)

df_np = data.frame(marker_level =  ests_np$cr$s,
                      controlled_risk =  ests_np$cr$est, 
                      ci_lower =  ests_np$cr$ci_lower, 
                      ci_upper =  ests_np$cr$ci_upper) 
write.csv(df_np, paste0( './source_data/Fig3_VIPP_',  mk, '.csv'))

mk_dist = dt_serum_PP_VIPP %>%
  filter(ph2.m1.sera == 1) %>%
  select(!!sym(mk), wt.m1.sera)

p = plot_cr2(ests_cox, ests_np, mk_dist, overall_incidence,
                     xlab_name = controlled_ve_names_vec[i],
                     title_name = 'Marker-Specific Risk',
                     t0 = 165, x_range = c(-0.5, 7.5),
                     x_breaks = seq(0, 7, 1),
                     x_labels =  c(expression(10^0), expression(10^1), expression(10^2),
                                   expression(10^3), expression(10^4), expression(10^5),
                                   expression(10^6), expression(10^7)))
p = p + geom_line(linewidth = 1.1) + 
  geom_ribbon(ggplot2::aes(ymin = ci_lower, ymax = ci_upper), 
              alpha = 0.2, linewidth = 1.1, linetype = "dotted") +
  theme_bw(base_size = 30) + 
  theme(legend.position = 'none') + ggtitle('')

saveRDS(p, paste0('./raw_data_for_figure/', controlled_ve_mk_vec[i], '_VIPP.RData'))

ggsave(paste0('./Figure/', controlled_ve_mk_vec[i], '_VIPP.pdf'),
       p, width = 12, height = 8, units = 'in')
}
```
