---
from: CoVPN 3008 
to: CoVPN 3008 Leadership
date: "`r format(Sys.Date(), '%B %d, %Y')`"
re:CoVPN 3008 Demographics table of Per-protocol Serum Immunogenicity Analysis Set participants
output:
  pdf_document:
    template: template_nopgnum.tex
    latex_engine: xelatex
  word_document:
    reference_docx: ../office_templates/word-pt-styles-reference-02.docx
  html_document:
    default
urlcolor: blue
bibliography: bibliography.bib
---





```{r knitr, echo = FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(knitr)
  library(kableExtra)
  library(hvtnReports)
  library(dplyr)
  library(tidyr) # for gather/spread
  library(ggplot2)
  library(broom) # for test.chisq tidying in chunk comparisons
  library(VISCfunctions) # for comparison functions
  library(haven) 
  library(rtf)
  library(stringr)
#  library(qdapRegex)
  library(gtsummary)
  library(conflicted)
  library(purrr)
})

fig_dev <- c("pdf", "png") # Create both pdf and png versions of each figure

# Determine output document type.
doctype <- opts_knit$get('rmarkdown.pandoc.to')
opts_chunk$set(echo = doctype == "html", message = doctype == "html",
               error = doctype == "html", warning = doctype == "html",
               # fig.width & fig.height control output file dimension
               # out.width & out.height control size in document
               fig.width = 10, fig.height = 7.5,
               out.width = "10in", out.height = "7.5in",
               dev = fig_dev, fig.path="../graph/cohort/")

# kable options - replace NA with 0
opts <- options(knitr.kable.NA = "0")

conflicts_prefer(dplyr::filter)

```

<!---setup table header name and footer parameter--->

```{r params, eval = doctype != "docx"}
   
tlf <- 
  list( 
  
    cohort_t = 
    c(table_name = "Cohort overall",
      table_header = "Participant characteristics by enrollment group, all countries",
      table_footer = "Any unknown, undefined, or missing values are due to incomplete live CRF data. Any levels not yet observed are not shown.")
    
  )


```

<!---setup table parameter--->

```{r param_setup, eval = doctype != "docx"}

### Protocol-specific parameters ---


# Do the same for the header
#header <- paste(header, collapse = "\\\\")

# Set table and figure numbers.
# Set the maximum number of rows to print, keeps tables from running off page.
tbl_num <- 1
fig_num <- 1
tbl.font <- 10

### Assay-specific parameters ---
subjid_var <- "ptid"




```

<!---load dataset--->

```{r dataio, eval = doctype != "docx"}

ds <- read.csv("../adata/COVID_realdata_ics_bama_nab_bcp20241210_processed_with_riskscore.csv") %>%
  mutate( VACCINE_f = case_when(AGRP2N == 2.1 | AGRP2N == 4.1 ~ "Hybrid",
                                   AGRP2N == 1 | AGRP2N == 3 | AGRP2N == 2.2 | AGRP2N == 4.2 ~ "Vaccine")) %>%
  select(-HIVVLBL2, -CD4BL2 ) # use the HIV and CD4count directly from verified ADSL

adsl <- read.csv("/trials/covpn/p3008/analysis/verification/adata/preM6/21APR2023/ADSL.csv")


table_dat <- adsl %>%
  mutate(AGRP2N = as.numeric(AGRP2N)) %>%
  # filter(FASFL=="Y", !is.na(AGRP2N)) %>%
  mutate(AGRP2N_f = case_when(AGRP2N==1 ~ "AG1",
                              AGRP2N==2.1 ~ "AG2-1",
                              AGRP2N==2.2 ~ "AG2-2",
                              AGRP2N==3 ~ "AG3",
                              AGRP2N==4.1 ~ "AG4-1",
                              AGRP2N==4.2 ~ "AG4-2"), 
         AGRP2N_f = factor(AGRP2N_f, levels=c("AG1","AG2-2","AG2-1","AG3","AG4-2","AG4-1")), 
         PCRBL_f = ifelse(PCRBL=="", "Missing", PCRBL),
         PCRBL_f = factor(PCRBL_f, levels=c("Negative", "Positive", "Missing")),
         CTRBL_f = ifelse(CTRBL=="", "Missing", CTRBL),
         CTRBL_f = factor(CTRBL_f, levels=c("Negative", "Positive", "Missing")),
         SEX_f = factor(SEX, levels=c("Male", "Female")),
         AGEGR1_f = paste(AGEGR1, "years"),
         AGEGR1_f = factor(AGEGR1_f, levels=c("<=40 years", ">40 years")),
         # receive vaccine doses as assigned in pre-Month 6 stage
         CORRECTD=if_else(TRT01P==TRT01A & TRT02P==TRT02A, "Yes", "No"),
         CORRECTD=factor(CORRECTD, levels=c("Yes", "No")),
         # categorization
         CD4BL_f = case_when(is.na(CD4BL) ~ "Missing",
                             CD4BL<200 ~ "<200",
                             CD4BL>=200 & CD4BL<350 ~ "200 - <350",
                             CD4BL>=350 & CD4BL<500 ~ "350 - <500",
                             CD4BL>=500 ~ ">=500"),
         CD4BL_f = factor(CD4BL_f, levels=c("<200", "200 - <350", "350 - <500", ">=500", "Missing")),
         HIVVLBL_f = case_when(is.na(HIVVLBL) & HIVDETBL=="" ~ "Missing",
                               HIVVLBL<50 | HIVDETBL=="Not detected" ~ "<50", # either viral load < 50 or not detected
                               HIVVLBL>=50 ~ ">=50"),
         HIVVLBL_f = factor(HIVVLBL_f, levels=c("<50", ">=50", "Missing")),
         ARTBLFL_f = case_when(ARTBLFL=="Y" ~ "On ART",
                             ARTBLFL=="N" ~ "Not on ART",
                             ARTBLFL=="" ~ "Missing"),
         ARTBLFL_f = factor(ARTBLFL_f, levels=c("On ART", "Not on ART", "Missing")),
         TB_active = case_when(TBFL=="Y" & TBTYP=="Active" ~ "Yes",
                               TBFL=="Y" & TBTYP=="Previous" ~ "No",
                               TBFL=="N" ~ "No",
                               TBFL=="" ~ "Missing"),
         TB_active = factor(TB_active, levels=c("Yes", "No", "Missing")),
         TB_previous = case_when(TBFL=="Y" & TBTYP=="Previous" ~ "Yes",
                                 TBFL=="Y" & TBTYP=="Active" ~ "No",
                                 TBFL=="N" ~ "No",
                                 TBFL=="" ~ "Missing"),
         TB_previous = factor(TB_previous, levels=c("Yes", "No", "Missing")),
         Subjectid = USUBJID )


 ds <- ds %>% full_join(., table_dat)


```




# Results
\clearpage

\blandscape <!-- This creates a landscaped page. \elandscape must follow somewhere -->
\tlf

```{cohort_tbl_seru, eval = doctype != "docx"}


 ds.seru <- ds %>%
  filter(ph2.m1.sera == 1 & IMMFL == "Y") 


# create table 1 with gtsummary package for study groups 1-4 categories
gtsummary_table_dat_grp1to4 <- ds.seru %>%
  select(VACCINE_f, CORRECTD, PCRBL_f, CTRBL_f, CTRYNAME, SEX_f, RACE,AGE, AGEGR1_f, TB_active, TB_previous) %>%
  tbl_summary(by=VACCINE_f,
              missing_text = "(Missing)",
              statistic = list("AGE" ~ "{median} ({min}, {max})"),
              digits = list(all_continuous() ~ c(1,1,1), # digits for continuous variables: mean, min, max
                            all_categorical() ~ c(0,1)), # digits for categorical variables
              label = list(CORRECTD ~ "Received vaccine doses as assigned",
                           PCRBL_f ~ "SARS-CoV-2 NAAT result - N (%)",
                           CTRBL_f ~ "Central lab anti-N test result - N (%)",
                           CTRYNAME ~"Country",
                           SEX_f ~ "Sex assigned at birth - N (%)",
                           RACE ~ "Race - N (%)",
                           AGE ~ "Median age (range) - years",
                           AGEGR1_f ~ "Age category - N (%)",
                           TB_active = "Active tuberculosis",
                           TB_previous = "History of tuberculosis")) %>%
  add_overall(last = TRUE) %>% # logical indicator to display overall column last in table
  as_tibble()

# create table 1 with gtsummary package for study groups 1-2 categories
gtsummary_table_dat_grp1to2 <- ds.seru %>%
 # filter(STRATARN %in% c(1,2)) %>% 
  select(VACCINE_f, CD4BL_f, CD4BL, HIVVLBL_f, HIVVLBL, ARTBLFL_f) %>%
  tbl_summary(by=VACCINE_f,
              missing_text = "(Missing)",
              statistic = list(c("CD4BL", "HIVVLBL") ~ "{median} ({p25}, {p75})"),
              digits = list(all_continuous() ~ c(1,1,1), # digits for continuous variables: median, p25, p75
                            all_categorical() ~ c(0,1)), # digits for categorical variables
              label = list(CD4BL ~ "Median CD4 count** (IQR) - cells/mm3",
                           CD4BL_f ~ "CD4 count** (cells/mm3) - N (%)",
                           HIVVLBL ~ "Median HIV-1 viral load** (IQR) - copies/mL",
                           HIVVLBL_f ~ "HIV viral load** (copies/mL) - N (%)",
                           ARTBLFL_f = "ART status**")) %>%
  add_overall(last = TRUE) %>% # logical indicator to display overall column last in table
  as_tibble()

# combine all table 1 categories
gtsummary_table_dat <- bind_rows(gtsummary_table_dat_grp1to4,
                                 gtsummary_table_dat_grp1to2) 

# kable table for output in PDF
kable_table_dat <- gtsummary_table_dat %>%
  # mutate_at(c("**Hybrid**, N = 138",  "**Vaccine**, N = 61",  "**Overall**, N = 199"), list(~ ifelse(row_number() %in% c(26:30, 31:32, 34:36, 37:38, 40:42), "-", .))) %>%
  # remove "(Missing)" row for continuous variables
  filter(`**Characteristic**`!="(Missing)") %>%
  mutate_all(list(~ gsub('%', '%', ., fixed=TRUE))) %>%
  mutate_all(list(~ replace_na(., "")))
names(kable_table_dat) <- gsub("Overall", "Total", gsub("\\*", "", names(kable_table_dat))) 

# add info rows
# info_row1 <- c("Point-of-care anti-Spike test result","Neg","Neg","Pos","Neg","Neg","Pos","")
# info_row2 <- c("Number of primary series vaccine doses assigned","2","2","1","2","2","1", "")
# info_row3 <- c("Overall SARS-CoV-2 status*","Neg","Pos","Pos","Neg","Pos","Pos","")
# kable_table_dat_all <- rbind(info_row1,
#                              info_row2,
#                              kable_table_dat[1:9,],
#                              info_row3,
#                              kable_table_dat[10:nrow(kable_table_dat),]) 

# output table
kable_table_dat %>% 
  kable(caption = "", 
      col.names = linebreak(names(kable_table_dat)),
      row.names = FALSE,
      booktabs = TRUE,
      linesep = "",
      longtable = TRUE,
      escape = FALSE,
      align = "l") %>%
  add_indent(positions = which(!(kable_table_dat$Characteristic %in% c(    "Received vaccine doses as assigned",
                                                                           "SARS-CoV-2 NAAT result - N (%)",
                                                                           "Central lab anti-N test result - N (%)",
                                                                           "Country",
                                                                           "Sex assigned at birth - N (%)",
                                                                           "Race - N (%)",
                                                                           "Median age (range) - years",
                                                                           "Age category - N (%)",
                                                                           "Active tuberculosis",
                                                                           "History of tuberculosis",
                                                                           "CD4 count** (cells/mm3) - N (%)",
                                                                           "Median CD4 count** (IQR) - cells/mm3",
                                                                           "HIV viral load** (copies/mL) - N (%)",
                                                                           "Median HIV-1 viral load** (IQR) - copies/mL",
                                                                           "ART status**")))) %>%
  kable_styling(latex_options = "repeat_header") %>%
  column_spec(1, width = "6.3cm") %>%
  column_spec(2:4, width = "2.2cm") %>%
  footnote(general_title = "",
            general = c(""),
           threeparttable=TRUE) %>%
  landscape


 ## report table in RTF format
rtf.w <- 11
 rtf.h <- 8.5
 rtf.fs <- 11
 rtf.mar <- c(1,1,1,1)
 rtf.na <- '0' 
 out.rtf <- "../report/cohort_table_Serum.doc"
 rtf <- RTF(out.rtf, 
            width = rtf.w, height = rtf.h, 
            font.size = rtf.fs, omi = rtf.mar)
 tbl_lab <- paste0("Table 1. Demographics of Per-protocol Serum Immunogenicity Analysis Set participants in the Hybrid Group, Vaccine Group, and Overall. ")
 
 rtf.tbl <- kable_table_dat 
 
  addHeader(rtf, tbl_lab,
          font.size = 12)
addTable(rtf, rtf.tbl,col.justify = c("L","L","C","C"))
addPageBreak(rtf, 
             width = rtf.w, height = rtf.h, 
             font.size = rtf.fs, omi = rtf.mar)
rtf::done(rtf)



```



```{cohort_tbl_ics, eval = doctype != "docx"}

 ds.ics <- ds %>%
  filter(ph2.m1.ics.cohort == 1 & IMMFL == "Y") 


# create table 1 with gtsummary package for study groups 1-4 categories
gtsummary_table_dat_grp1to4 <- ds.ics %>%
  select(VACCINE_f, CORRECTD, PCRBL_f, CTRBL_f, CTRYNAME, SEX_f, RACE,AGE, AGEGR1_f, TB_active, TB_previous) %>%
  tbl_summary(by=VACCINE_f,
              missing_text = "(Missing)",
              statistic = list("AGE" ~ "{median} ({min}, {max})"),
              digits = list(all_continuous() ~ c(1,1,1), # digits for continuous variables: mean, min, max
                            all_categorical() ~ c(0,1)), # digits for categorical variables
              label = list(CORRECTD ~ "Received vaccine doses as assigned",
                           PCRBL_f ~ "SARS-CoV-2 NAAT result - N (%)",
                           CTRBL_f ~ "Central lab anti-N test result - N (%)",
                           CTRYNAME ~"Country",
                           SEX_f ~ "Sex assigned at birth - N (%)",
                           RACE ~ "Race - N (%)",
                           AGE ~ "Median age (range) - years",
                           AGEGR1_f ~ "Age category - N (%)",
                           TB_active = "Active tuberculosis",
                           TB_previous = "History of tuberculosis")) %>%
  add_overall(last = TRUE) %>% # logical indicator to display overall column last in table
  as_tibble()

# create table 1 with gtsummary package for study groups 1-2 categories
gtsummary_table_dat_grp1to2 <- ds.ics %>%
 # filter(STRATARN %in% c(1,2)) %>% 
  select(VACCINE_f, CD4BL_f, CD4BL, HIVVLBL_f, HIVVLBL, ARTBLFL_f) %>%
  tbl_summary(by=VACCINE_f,
              missing_text = "(Missing)",
              statistic = list(c("CD4BL", "HIVVLBL") ~ "{median} ({p25}, {p75})"),
              digits = list(all_continuous() ~ c(1,1,1), # digits for continuous variables: median, p25, p75
                            all_categorical() ~ c(0,1)), # digits for categorical variables
              label = list(CD4BL ~ "Median CD4 count** (IQR) - cells/mm3",
                           CD4BL_f ~ "CD4 count** (cells/mm3) - N (%)",
                           HIVVLBL ~ "Median HIV-1 viral load** (IQR) - copies/mL",
                           HIVVLBL_f ~ "HIV viral load** (copies/mL) - N (%)",
                           ARTBLFL_f = "ART status**")) %>%
  add_overall(last = TRUE) %>% # logical indicator to display overall column last in table
  as_tibble()

# combine all table 1 categories
gtsummary_table_dat <- bind_rows(gtsummary_table_dat_grp1to4,
                                 gtsummary_table_dat_grp1to2) 

# kable table for output in PDF
kable_table_dat <- gtsummary_table_dat %>%
  # remove "(Missing)" row for continuous variables
  filter(`**Characteristic**`!="(Missing)") %>%
  mutate_all(list(~ gsub('%', '%', ., fixed=TRUE))) %>%
  mutate_all(list(~ replace_na(., "")))
names(kable_table_dat) <- gsub("Overall", "Total", gsub("\\*", "", names(kable_table_dat))) 

# add info rows
# info_row1 <- c("Point-of-care anti-Spike test result","Neg","Neg","Pos","Neg","Neg","Pos","")
# info_row2 <- c("Number of primary series vaccine doses assigned","2","2","1","2","2","1", "")
# info_row3 <- c("Overall SARS-CoV-2 status*","Neg","Pos","Pos","Neg","Pos","Pos","")
# kable_table_dat_all <- rbind(info_row1,
#                              info_row2,
#                              kable_table_dat[1:9,],
#                              info_row3,
#                              kable_table_dat[10:nrow(kable_table_dat),]) 

# output table
kable_table_dat %>% 
  kable(caption = "", 
      col.names = linebreak(names(kable_table_dat)),
      row.names = FALSE,
      booktabs = TRUE,
      linesep = "",
      longtable = TRUE,
      escape = FALSE,
      align = "l") %>%
  add_indent(positions = which(!(kable_table_dat$Characteristic %in% c(    "Received vaccine doses as assigned",
                                                                           "SARS-CoV-2 NAAT result - N (%)",
                                                                           "Central lab anti-N test result - N (%)",
                                                                           "Country",
                                                                           "Sex assigned at birth - N (%)",
                                                                           "Race - N (%)",
                                                                           "Median age (range) - years",
                                                                           "Age category - N (%)",
                                                                           "Active tuberculosis",
                                                                           "History of tuberculosis",
                                                                           "CD4 count** (cells/mm3) - N (%)",
                                                                           "Median CD4 count** (IQR) - cells/mm3",
                                                                           "HIV viral load** (copies/mL) - N (%)",
                                                                           "Median HIV-1 viral load** (IQR) - copies/mL",
                                                                           "ART status**")))) %>%
#  add_header_above(c(" ", "Study Group 1" = 2, "Study Group 2" = 1, "Study Group 3" = 2, "Study Group 4" = 1, " ")) %>%
  kable_styling(latex_options = "repeat_header") %>%
  column_spec(1, width = "6.3cm") %>%
  column_spec(2:4, width = "2.2cm") %>%
  footnote(general_title = "",
            general = c(""),
           threeparttable=TRUE) %>%
  landscape


 ## report table in RTF format
rtf.w <- 11
 rtf.h <- 8.5
 rtf.fs <- 11
 rtf.mar <- c(1,1,1,1)
 rtf.na <- '0' 
 out.rtf <- "../report/cohort_table_ICS_RIS.doc"
 rtf <- RTF(out.rtf, 
            width = rtf.w, height = rtf.h, 
            font.size = rtf.fs, omi = rtf.mar)
 tbl_lab <- paste0("Table 1. Demographics of Random Immunogenicity Subset ICS Immunogenicity Analysis Set participants in the Hybrid Group, Vaccine Group, and Overall. ")
 
 rtf.tbl <- kable_table_dat 
 
  addHeader(rtf, tbl_lab,
          font.size = 12)
addTable(rtf, rtf.tbl,col.justify = c("L","L","C","C"))
addPageBreak(rtf, 
             width = rtf.w, height = rtf.h, 
             font.size = rtf.fs, omi = rtf.mar)
rtf::done(rtf)


```



```{cohort_tbl_ics_sub, eval = doctype != "docx"}

 ds.ics <- ds %>%
  filter(ph2.m1.ics.cohort == 1 & IMMFL == "Y" & AGRP2N %in% c(1, 2.1)) 


# create table 1 with gtsummary package for study groups 1-4 categories
gtsummary_table_dat_grp1to4 <- ds.ics %>%
  select(VACCINE_f, CORRECTD, PCRBL_f, CTRBL_f, CTRYNAME, SEX_f, RACE,AGE, AGEGR1_f, TB_active, TB_previous) %>%
  tbl_summary(by=VACCINE_f,
              missing_text = "(Missing)",
              statistic = list("AGE" ~ "{median} ({min}, {max})"),
              digits = list(all_continuous() ~ c(1,1,1), # digits for continuous variables: mean, min, max
                            all_categorical() ~ c(0,1)), # digits for categorical variables
              label = list(CORRECTD ~ "Received vaccine doses as assigned",
                           PCRBL_f ~ "SARS-CoV-2 NAAT result - N (%)",
                           CTRBL_f ~ "Central lab anti-N test result - N (%)",
                           CTRYNAME ~"Country",
                           SEX_f ~ "Sex assigned at birth - N (%)",
                           RACE ~ "Race - N (%)",
                           AGE ~ "Median age (range) - years",
                           AGEGR1_f ~ "Age category - N (%)",
                           TB_active = "Active tuberculosis",
                           TB_previous = "History of tuberculosis")) %>%
  add_overall(last = TRUE) %>% # logical indicator to display overall column last in table
  as_tibble()

# create table 1 with gtsummary package for study groups 1-2 categories
gtsummary_table_dat_grp1to2 <- ds.ics %>%
 # filter(STRATARN %in% c(1,2)) %>% 
  select(VACCINE_f, CD4BL_f, CD4BL, HIVVLBL_f, HIVVLBL, ARTBLFL_f) %>%
  tbl_summary(by=VACCINE_f,
              missing_text = "(Missing)",
              statistic = list(c("CD4BL", "HIVVLBL") ~ "{median} ({p25}, {p75})"),
              digits = list(all_continuous() ~ c(1,1,1), # digits for continuous variables: median, p25, p75
                            all_categorical() ~ c(0,1)), # digits for categorical variables
              label = list(CD4BL ~ "Median CD4 count** (IQR) - cells/mm3",
                           CD4BL_f ~ "CD4 count** (cells/mm3) - N (%)",
                           HIVVLBL ~ "Median HIV-1 viral load** (IQR) - copies/mL",
                           HIVVLBL_f ~ "HIV viral load** (copies/mL) - N (%)",
                           ARTBLFL_f = "ART status**")) %>%
  add_overall(last = TRUE) %>% # logical indicator to display overall column last in table
  as_tibble()

# combine all table 1 categories
gtsummary_table_dat <- bind_rows(gtsummary_table_dat_grp1to4,
                                 gtsummary_table_dat_grp1to2) 

# kable table for output in PDF
kable_table_dat <- gtsummary_table_dat %>%
  # remove "(Missing)" row for continuous variables
  filter(`**Characteristic**`!="(Missing)") %>%
  mutate_all(list(~ gsub('%', '%', ., fixed=TRUE))) %>%
  mutate_all(list(~ replace_na(., "")))
names(kable_table_dat) <- gsub("Overall", "Total", gsub("\\*", "", names(kable_table_dat))) 

# add info rows
# info_row1 <- c("Point-of-care anti-Spike test result","Neg","Neg","Pos","Neg","Neg","Pos","")
# info_row2 <- c("Number of primary series vaccine doses assigned","2","2","1","2","2","1", "")
# info_row3 <- c("Overall SARS-CoV-2 status*","Neg","Pos","Pos","Neg","Pos","Pos","")
# kable_table_dat_all <- rbind(info_row1,
#                              info_row2,
#                              kable_table_dat[1:9,],
#                              info_row3,
#                              kable_table_dat[10:nrow(kable_table_dat),]) 

# output table
kable_table_dat %>% 
  kable(caption = "", 
      col.names = linebreak(names(kable_table_dat)),
      row.names = FALSE,
      booktabs = TRUE,
      linesep = "",
      longtable = TRUE,
      escape = FALSE,
      align = "l") %>%
  add_indent(positions = which(!(kable_table_dat$Characteristic %in% c(    "Received vaccine doses as assigned",
                                                                           "SARS-CoV-2 NAAT result - N (%)",
                                                                           "Central lab anti-N test result - N (%)",
                                                                           "Country",
                                                                           "Sex assigned at birth - N (%)",
                                                                           "Race - N (%)",
                                                                           "Median age (range) - years",
                                                                           "Age category - N (%)",
                                                                           "Active tuberculosis",
                                                                           "History of tuberculosis",
                                                                           "CD4 count** (cells/mm3) - N (%)",
                                                                           "Median CD4 count** (IQR) - cells/mm3",
                                                                           "HIV viral load** (copies/mL) - N (%)",
                                                                           "Median HIV-1 viral load** (IQR) - copies/mL",
                                                                           "ART status**")))) %>%
#  add_header_above(c(" ", "Study Group 1" = 2, "Study Group 2" = 1, "Study Group 3" = 2, "Study Group 4" = 1, " ")) %>%
  kable_styling(latex_options = "repeat_header") %>%
  column_spec(1, width = "6.3cm") %>%
  column_spec(2:4, width = "2.2cm") %>%
  footnote(general_title = "",
            general = c(""),
           threeparttable=TRUE) %>%
  landscape


 ## report table in RTF format
rtf.w <- 11
 rtf.h <- 8.5
 rtf.fs <- 11
 rtf.mar <- c(1,1,1,1)
 rtf.na <- '0' 
 out.rtf <- "../report/cohort_table_ics_RIS_PLWH.doc"
 rtf <- RTF(out.rtf, 
            width = rtf.w, height = rtf.h, 
            font.size = rtf.fs, omi = rtf.mar)
 tbl_lab <- paste0("Table 1. Demographics of Random Immunogenicity Analysis Set for participants living with HIV in the Hybrid Group, Vaccine Group, and Overall. ")
 
 rtf.tbl <- kable_table_dat 
 
  addHeader(rtf, tbl_lab,
          font.size = 12)
addTable(rtf, rtf.tbl,col.justify = c("L","L","C","C"))
addPageBreak(rtf, 
             width = rtf.w, height = rtf.h, 
             font.size = rtf.fs, omi = rtf.mar)
rtf::done(rtf)
```



```{tbl1_serum, eval = doctype != "docx"}

source("../macro/make_functions.R")

#### calculate the IgG N Index (AU/ml) (no response call) , IgG Spike BA.5 (AU/ml), nAb ID50 BA.4/5 (AU/ml)response rate and geometric mean by case and non case
### read BAMA response varibal

 ds.seru <- ds %>%
  filter(ph2.m1.sera == 1 ) 


bama.resp.a23 <-  read.csv("../pdata/covpn3008_bammsd_20240206/pdata/CoVPN3008_BAMMSD_20240405_IQC_IgG.csv") %>%
  filter(analyte %in% c("SARS-CoV-2 Spike (BA.5)")  & aval_cd == "response") %>%
  filter(visitno %in% c(2, 3, 4)) %>%
  mutate(Subjectid = paste0("C3008-", ptid),
         visitno = case_when(visitno == 2 ~ "response_M0",
                             visitno ==3 |visitno ==4 ~ "response_M1")) %>%
  select("Subjectid", "visitno", "analyte", "aval") %>%
  pivot_wider(names_from = visitno, values_from = aval) %>%
  mutate(antigen.lab = "A23" )

bama.au.a23 <-  read.csv("../adata/COVID_bama_resp.csv") %>%
  filter(antigen.lab == "A23" ) %>%
    filter(visitno %in% c(2, 3, 4)) %>%
  mutate(Subjectid = paste0("C3008-", ptid),
         visitno = case_when(visitno == 2 ~ "obsconc_AU_tuc_M0",
                             visitno ==3 |visitno ==4 ~ "obsconc_AU_tuc_M1")) %>%
  select(Subjectid, antigen.lab, obsconc_AU_tuc, isotype, concentration_units,visitno, pos_cut) %>%
  pivot_wider(names_from = visitno, values_from = obsconc_AU_tuc)


bama.a23 <- bama.resp.a23 %>% full_join(bama.au.a23) %>%
  select(-analyte)


bama.a15 <- read.csv("../adata/COVID_bama_resp.csv") %>%
  select(ptid, antigen.lab, obsconc_AU_tuc, isotype, concentration_units,visitno, pos_cut)%>% 
   filter(visitno %in% c(2, 3, 4)) %>%
  rename(Subjectid = ptid) %>%
  mutate(Subjectid = paste0("C3008-", as.character(Subjectid)),
         visitno = case_when(visitno == 2 ~ "M0",
                             visitno ==3 |visitno ==4 ~ "M1"))%>%
  filter(antigen.lab == "A15" ) 


###### make response call by the following steps
###### 1. [for baseline visit 2]For all antigen expect IgG N, keep using the positivity call offered by lab. 
###### 2. [for baseline visit 2]For IgG N, if AU < 3970/2, the response is negative, if AU >= 3970/2, the response is positive
bama.a15 <- bama.a15 %>% 
  mutate(response = case_when(obsconc_AU_tuc < pos_cut ~ 0,
                              obsconc_AU_tuc >= pos_cut ~ 1)) %>%
  pivot_wider(names_from = visitno, values_from = c(response, obsconc_AU_tuc)) 

bama.resp <- rbind(bama.a15, bama.a23)

##### 3.If the M0 = neg, Peak resp = 1 (response by lab) ~ resp =1 at Peak
##### If the M0 = neg, Peak resp = 0 (by lab) ~ resp = 0 at Peak
##### If M0 = pos and Peak resp = 1 and need to 4 folder increase ~ resp at Peak =1
##### If M0 = pos ~ Peak resp = 1 and change less 4 fold ~ resp at Peak = 0 

bama.resp <- bama.resp %>%
  mutate(au_ratio = obsconc_AU_tuc_M1/obsconc_AU_tuc_M0) %>%
  mutate(response_M1_adj = case_when(response_M0 == 0 & response_M1 == 0 ~ 0,
                                     response_M0 == 0 & response_M1 == 1 ~ 1,
                                     response_M0 == 1 & response_M1 == 0 ~ 0,
                                     response_M0 == 1 & response_M1 == 1 & au_ratio >= 4 ~ 1,
                                     response_M0 == 1 & response_M1 == 1 & au_ratio < 4 ~ 0)) %>%
  select(Subjectid, antigen.lab, response_M0, response_M1_adj) %>%
  pivot_wider(names_from = antigen.lab, values_from = c(response_M0, response_M1_adj)) %>%
  rename(BA23AU.resp  = response_M0_A23,
         M1A23AU.resp = response_M1_adj_A23,
         BA15AU.resp  = response_M0_A15,
         M1A15AU.resp = response_M1_adj_A15)



#### read nab data 

nab.resp <- read.csv("../pdata/COVPN3008_NAB07_SA_20240705/pdata/COVPN3008_NAB07_SA_resp_long_20240711.csv") %>%
  filter(isolate == "BA.4/5" & aval_cd == "response_ID50" & visitno %in% c(2,3,4)) %>%
  mutate(Subjectid = paste0("C3008-", ptid),
         visitno = case_when(visitno == 2 ~ "BBA45ID50.resp",
                             visitno ==3 |visitno ==4 ~ "M1BA45ID50.resp")) %>%
  select("Subjectid", "visitno", "aval") %>%
  pivot_wider(names_from = visitno, values_from = aval)


ds.tmp.seru <- left_join(ds.seru %>%
                      mutate(EventIndPrimary1 = case_when(EventIndPrimary1 == 0 ~ "Non-cases",
                                                           EventIndPrimary1 == 1 ~ "Cases")) %>%
  select("Subjectid", "BIGGMSDA15AU", "BIGGMSDA23AU", "M1IGGMSDA15AU", "M1IGGMSDA23AU", "BBA45ID50", "M1BA45ID50", "EventIndPrimary1", "VACCINE_f"),
  bama.resp) %>%
  left_join(.,nab.resp) %>%
  mutate(BA23AU.resp = as.numeric(BA23AU.resp),
         M1A23AU.resp = as.numeric(M1A23AU.resp),
         BA15AU.resp = as.numeric(BA15AU.resp),
         M1A15AU.resp = as.numeric(M1A15AU.resp),
         BBA45ID50.resp = as.numeric(BBA45ID50.resp),
         M1BA45ID50.resp = as.numeric(M1BA45ID50.resp))


#### generate response rate table
tbl <- data_frame()

for(resp in c("BA15AU.resp", "BA23AU.resp", "BBA45ID50.resp", "M1A15AU.resp", "M1A23AU.resp", "M1BA45ID50.resp")){
  
   tbl.tmp  <- point_response(data = ds.tmp.seru , response_var = resp , group_var = c("EventIndPrimary1", "VACCINE_f"), out_names = c("Response Rate", "95% CI") ) %>%
     mutate(antigen = resp)
  
   tbl <- rbind(tbl, tbl.tmp)
  
}

tbl1.1 <- tbl %>% pivot_wider(names_from = EventIndPrimary1, values_from = c( `Response Rate`, `95% CI`) ) %>%
  mutate(resp.diff = paste0( round(as.numeric(qdapRegex::ex_between(`Response Rate_Non-cases`, "= ", "%")) - as.numeric(qdapRegex::ex_between(`Response Rate_Cases`, "= ", "%")), 1), "%") ) %>%
  mutate(resp.diff = case_when(resp.diff == "NA%" ~ "",
                               TRUE ~ resp.diff)) %>%
  mutate(antigen = case_when(antigen == "BA15AU.resp" ~ "BIGGMSDA15AU",
                             antigen == "BA23AU.resp" ~ "BIGGMSDA23AU",
                             antigen == "M1A15AU.resp" ~ "M1IGGMSDA15AU",
                             antigen == "M1A23AU.resp" ~ "M1IGGMSDA23AU",
                             antigen == "BBA45ID50.resp" ~ "BBA45ID50",
                             antigen == "M1BA45ID50.resp" ~ "M1BA45ID50"))


#### generate magnitude table
tbl <- data_frame()

for(magnitude in c("BIGGMSDA15AU", "BIGGMSDA23AU", "M1IGGMSDA15AU", "M1IGGMSDA23AU", "BBA45ID50", "M1BA45ID50")){
  
   tbl.tmp <- desc_stats_table(data = ds.tmp.seru , response_var = magnitude , group_var = c("EventIndPrimary1", "VACCINE_f"), log =TRUE, ci =TRUE, na.rm=TRUE ) %>%
     mutate(antigen = magnitude,
            `Geometric Mean (GM)` = gm_mean.est,
            `GM 95% CI` = paste0("(", gm_mean.ci_l, ", " , gm_mean.ci_u, ")"))%>%
     select("EventIndPrimary1", "VACCINE_f", "n", `Geometric Mean (GM)`, `GM 95% CI`, "antigen")
  
   tbl <- rbind(tbl, tbl.tmp)
  
}

tbl1.2 <- tbl %>% pivot_wider(names_from = EventIndPrimary1, values_from = c( n, `Geometric Mean (GM)`, `GM 95% CI`) ) %>%
  mutate(`GM Ratio` =  round(`Geometric Mean (GM)_Non-cases`/`Geometric Mean (GM)_Cases`, 1)) %>%
 select(VACCINE_f, antigen, n_Cases,  `Geometric Mean (GM)_Cases`,  `GM 95% CI_Cases`, `n_Non-cases`, `Geometric Mean (GM)_Non-cases` , `GM 95% CI_Non-cases`)


ds.seru.resp.long <- ds.tmp.seru %>% 
  select(Subjectid, EventIndPrimary1, VACCINE_f, BA15AU.resp, BA23AU.resp, M1A15AU.resp, M1A23AU.resp, BBA45ID50.resp, M1BA45ID50.resp ) %>%
  pivot_longer(cols = contains("resp"),
               names_to = "assay",
               values_to = "response") %>%
  mutate(assay = case_when(assay == "BA15AU.resp" ~ "BIGGMSDA15AU",
                             assay == "BA23AU.resp" ~ "BIGGMSDA23AU",
                             assay == "M1A15AU.resp" ~ "M1IGGMSDA15AU",
                             assay == "M1A23AU.resp" ~ "M1IGGMSDA23AU",
                             assay == "BBA45ID50.resp" ~ "BBA45ID50",
                             assay == "M1BA45ID50.resp" ~ "M1BA45ID50"))
  
  
ds.seru.value.long <- ds.tmp.seru %>% 
  select(Subjectid, EventIndPrimary1, VACCINE_f, BIGGMSDA15AU, BIGGMSDA23AU, M1IGGMSDA15AU, M1IGGMSDA23AU, BBA45ID50, M1BA45ID50) %>%
  pivot_longer(cols = BIGGMSDA15AU:M1BA45ID50 ,
               names_to = "assay",
               values_to = "value") 
  
  
ds.seru.long <- ds.seru.resp.long %>% full_join(., ds.seru.value.long)%>%
  mutate(value = log10(value))


resp.diff.ci <- ds.seru.resp.long %>%
  group_by(assay , EventIndPrimary1, VACCINE_f) %>%
  summarize(n = n(),
            r = sum(response, na.rm=TRUE))%>% 
  pivot_wider(names_from = EventIndPrimary1, values_from = c( n, r))

  resp.ci <- data_frame()

  for(i in 1:nrow(resp.diff.ci)){
    
    ds.tmp <-resp.diff.ci[i,] 
    
    
    resp.ci.tmp <- ds.tmp %>%
  mutate(lcl = prop.test(x = c(ds.tmp$r_Cases, ds.tmp$`r_Non-cases`), 
                               n = c(ds.tmp$n_Cases, ds.tmp$`n_Non-cases`), 
                               alternative = "two.sided", 
                               correct = FALSE)$conf.int[1],
         
         ucl = prop.test(x = c(ds.tmp$r_Cases, ds.tmp$`r_Non-cases`), 
                               n = c(ds.tmp$n_Cases, ds.tmp$`n_Non-cases`), 
                               alternative = "two.sided", 
                               correct = FALSE)$conf.int[2]) 
    
    resp.ci <- rbind(resp.ci, resp.ci.tmp)
    
  }
  
  
  resp.ci <- resp.ci %>%
    mutate(lcl =round(lcl*100,2),
           ucl =round(ucl*100, 2)) %>%
    mutate(`resp CI` = paste0("(", lcl, "%", ", ", ucl, "%)")) %>%
    select(assay,  VACCINE_f, `resp CI`) %>%
    rename(antigen = assay)

  
  
  
  


# joining dataframes with regression estimates and model summaries
tbl1.3 <- dplyr::full_join(
 # to get a tidy dataframe of regression estimates
  x = ds.seru.long %>%
    group_by(VACCINE_f, assay) %>% 
    group_modify(.f = ~ broom::tidy(lm(value ~ EventIndPrimary1, data = .x), conf.int = TRUE)),
  # to get a tidy dataframe of model summaries
  y = ds.seru.long %>%
    group_by(VACCINE_f, assay) %>%
    group_modify(.f = ~ broom::glance(lm(value ~ EventIndPrimary1, data = .x))),
  by = c("VACCINE_f", "assay")
) %>%
  dplyr::ungroup(x = .) %>%
  filter(term == "EventIndPrimary1Non-cases") %>%
  mutate(`GM Ratio` = round(10^estimate, 2),
         `GM Ratio CI` = paste0("(", round(10^conf.low, 2), ", ", round(10^conf.high, 2), ")")) %>%
  select(VACCINE_f, assay, `GM Ratio`, `GM Ratio CI`) %>%
  rename(antigen = assay)



tbl1 <- full_join(tbl1.1, tbl1.2) %>%
  full_join(resp.ci) %>%
  full_join(., tbl1.3) %>%
  mutate(VACCINE_f = factor(VACCINE_f, levels = c("Hybrid", "Vaccine")),
         antigen = factor(antigen, levels = c("BIGGMSDA15AU", "BIGGMSDA23AU", "BBA45ID50", "M1IGGMSDA15AU", "M1IGGMSDA23AU", "M1BA45ID50"))) %>%
  arrange(VACCINE_f, antigen) %>%
  select(VACCINE_f, antigen, n_Cases, `Response Rate_Cases`, `95% CI_Cases`, `Geometric Mean (GM)_Cases`, `GM 95% CI_Cases`, `n_Non-cases`, `Response Rate_Non-cases`, `95% CI_Non-cases`,
         `Geometric Mean (GM)_Non-cases`, `GM 95% CI_Non-cases`,
          `resp.diff`, `resp CI`  , `GM Ratio`, `GM Ratio CI`)




## report table in RTF format
rtf.w <- 11
 rtf.h <- 8.5
 rtf.fs <- 11
 rtf.mar <- c(1,1,1,1)
 rtf.na <- '0' 
 out.rtf <- "../report/tbl1_serum.doc"
 rtf <- RTF(out.rtf, 
            width = rtf.w, height = rtf.h, 
            font.size = rtf.fs, omi = rtf.mar)
 tbl_lab <- paste0("Table 1. Immune marker response frequencies at M0 and vaccine-response frequencies at Peak (when applicable) by COVID-19 outcome status, geometric means (GMs) by COVID-19 outcome status, and their comparisons between non-cases and cases ")
 
 rtf.tbl <- tbl1 
 
  addHeader(rtf, tbl_lab,
          font.size = 12)
addTable(rtf, rtf.tbl,col.justify = c("L","L","C","C","C","C","C","C","C","C","C","C","C","C","C","C"))
addPageBreak(rtf, 
             width = rtf.w, height = rtf.h, 
             font.size = rtf.fs, omi = rtf.mar)
rtf::done(rtf)



```



```{tbl1_ics, eval = doctype != "docx"}

#### calculate the  CD4+ T-cells expressing IFNg and/or IL-2 BA.5 N,  CD4+ T-cells expressing IFNg and/or IL-2 and/or CD154 BA.5 N ,  CD4+ T-cells expressing IFNg and/or IL-2 BA.5 Spike,  CD4+ T-cells expressing IFNg and/or IL-2 and/or CD154 BA.5 Spike response rate and geometric mean by case and non case
### read ics data
### in the iCS data, keep IFNg_OR_IL2 (C03), IFNg_OR_IL2_OR_154-"Spike BA.4/5"(A11), "Spike Wuhan"(A13), "BA.4-5 N" (A02), "Wuhan N" (A08)

 ds.ics <- ds %>%
  filter(ph2.m1.ics == 1 ) 
# 
# %>%
#   select("Subjectid", "EventIndPrimary1", "VACCINE_f","AGRP2N", "BT4C03A11", "BT4C03A13", "BT4C03A02", "BT4C03A08",
#                                                                  "M1T4C03A11", "M1T4C03A13", "M1T4C03A02", "M1T4C03A08",
#                                                                   "BT8C03A11", "BT8C03A13", "BT8C03A02", "BT8C03A08",
#                                                                  "M1T8C03A11", "M1T8C03A13", "M1T8C03A02", "M1T8C03A08",
#                                                                   "BT4C04A11", "BT4C04A13", "BT4C04A02", "BT4C04A08",
#                                                                  "M1T4C04A11", "M1T4C04A13", "M1T4C04A02", "M1T4C04A08" )
# 
# ds.ics.long <- ds.ics %>%
#   pivot_longer(cols = BT4C03A11:M1T8C03A08,
#                names_to = "assay",
#                values_to = "pctpos_adj") %>%
#   mutate(visitno = case_when(startsWith(assay, "B") ~ "M0",
#                              startsWith(assay, "M1") ~ "M1")) %>%
#   mutate(tcell = case_when(grepl("T4", assay) ~ " CD4+",
#                            grepl("T8", assay,)  ~ "CD8+")) %>%
#   mutate(cytokine = case_when(grepl("C03", assay) ~ " IFNg_OR_IL2",
#                            grepl("T8", assay,)  ~ "CD8+"))
  


ics.pool <- read.csv("../pdata/covpn3008_fcm08_CH_AP59_20231120_case-cohort_correlates/pdata/covpn3008_fcm08_CH_AP59_pool_p.csv") %>%
  filter(cytokine %in% c('IFNg_OR_IL2', "IFNg_OR_IL2_OR_154") &
           antigen %in% c("Spike BA.4/5",
                          "Spike Wuhan"
                          ) &
           tcellsub == "4+")

ics.resp <-
  read.csv("../pdata/covpn3008_fcm08_CH_AP59_20231120_case-cohort_correlates/pdata/covpn3008_fcm08_CH_AP59_rpt_resp.csv") %>%
  filter(cytokine %in% c('IFNg_OR_IL2', "IFNg_OR_IL2_OR_154") &
           antigen %in% c("BA.4-5 N",
                          "Wuhan N")&
           tcellsub == "4+")



pubid <- read_sas("../adata/pubid.sas7bdat") %>%
  select(SUBJID, PUBID)%>%
  rename(ptid = SUBJID)

colnames(pubid) <- tolower(colnames(pubid))


colnames(ics.pool) <- tolower(colnames(ics.pool))
colnames(ics.resp) <- tolower(colnames(ics.resp))


ics <- bind_rows(
  ics.pool %>%
    dplyr::select(labid, ptid, visitno,
           tcellsub, cytokine, antigen, pctpos_adj, response),
  ics.resp %>% dplyr::select(labid, ptid, visitno,
                      tcellsub, cytokine, antigen, pctpos_adj, response)
  ) %>%
  mutate(Subjectid = paste0("C3008-", ptid)) %>%
  mutate(pctpos_adj = case_when(pctpos_adj <= 0.01 ~ 0.01,
                                TRUE ~ pctpos_adj))



ics.vacc <- read.csv("../pdata/ICS_responder.csv") %>%
  rename(visitno = type,
         cytokine = cyt,
         antigen = ant,
         VACCINE_f = gp,
         EventIndPrimary1 = status) %>%
  mutate(vacc.response = case_when(responder == "TRUE" ~ 1,
                               responder == "FALSE" ~ 0)) %>%
  left_join(., pubid) %>%
  select(ptid, vacc.response, visitno, cytokine, antigen, EventIndPrimary1, VACCINE_f) %>%
  mutate(antigen = case_when(antigen == "N_BA45" ~ "BA.4-5 N",
                             antigen == "Spike_BA45" ~ "Spike BA.4/5")) %>%
  mutate(EventIndPrimary1 = case_when(EventIndPrimary1 == "case" ~ "Cases",
                                      EventIndPrimary1 == "noncase"  ~ "Non-cases")) %>%
  mutate(visitno = case_when(visitno == "Peak" ~ "M1",
                             visitno == "M0"  ~ "M0")) %>%
  mutate(cytokine = case_when(cytokine == "IFNg_OR_IL2" ~ "IFNg_OR_IL2",
         cytokine == "IFNg_OR_IL2_154" ~ "IFNg_OR_IL2_OR_154")) %>%
   mutate(Subjectid = paste0("C3008-", ptid)) %>%
  select(-ptid)

         
         
ds.tmp.ics <- left_join(ds.ics %>%
                      mutate(EventIndPrimary1 = case_when(EventIndPrimary1 == 0 ~ "Non-cases",
                                                           EventIndPrimary1 == 1 ~ "Cases")) %>%
  select("Subjectid", "EventIndPrimary1", "VACCINE_f","AGRP2N") %>%
    filter(AGRP2N %in% c(1, 2.1)),
  ics) %>%
  mutate(visitno = case_when(visitno == 2 ~ "M0",
                             visitno == 3 | visitno == 4 ~ "M1")) %>%
  mutate(value = log10(pctpos_adj)) %>%
  left_join(., ics.vacc)


ds.tmp.ics.tbl1 <- ds.tmp.ics %>% filter(antigen %in% c("Spike BA.4/5", "BA.4-5 N"))


tbl2.1 <- point_response(data = ds.tmp.ics.tbl1, response_var = "vacc.response" , group_var = c("visitno", "cytokine", "antigen" ,"EventIndPrimary1", "VACCINE_f"), out_names = c("Response Rate", "95% CI") ) %>% 
  pivot_wider(names_from = EventIndPrimary1, values_from = c( `Response Rate`, `95% CI`) ) %>%
  mutate(resp.diff = paste0( round(as.numeric(qdapRegex::ex_between(`Response Rate_Non-cases`, "= ", "%")) - as.numeric(qdapRegex::ex_between(`Response Rate_Cases`, "= ", "%")), 1), "%") ) %>%
  mutate(resp.diff = case_when(resp.diff == "NA%" ~ "",
                               TRUE ~ resp.diff))


tbl2.2 <- desc_stats_table(data = ds.tmp.ics.tbl1, response_var = "pctpos_adj" , group_var = c("visitno", "cytokine", "antigen" ,"EventIndPrimary1", "VACCINE_f"), log =TRUE, ci =TRUE, na.rm=TRUE,digits = 4 )%>% 
     mutate(`Geometric Mean (GM)` = gm_mean.est,
            `GM 95% CI` = case_when(n == 1 ~ "-",
                          TRUE ~ paste0("(", gm_mean.ci_l, ", " , gm_mean.ci_u, ")"))) %>%
     select("visitno", "cytokine", "antigen" , "EventIndPrimary1", "VACCINE_f", "n", `Geometric Mean (GM)`, `GM 95% CI`) %>%
  pivot_wider(names_from = EventIndPrimary1, values_from = c( n, `Geometric Mean (GM)`, `GM 95% CI`) ) %>%
 select(VACCINE_f, visitno, cytokine, antigen, `n_Cases`, `Geometric Mean (GM)_Cases` , `GM 95% CI_Cases`, `n_Non-cases`, `Geometric Mean (GM)_Non-cases` , `GM 95% CI_Non-cases`)

 

# joining dataframes with regression estimates and model summaries
tbl2.3 <- dplyr::full_join(
 # to get a tidy dataframe of regression estimates
  x = ds.tmp.ics.tbl1 %>%
    group_by(VACCINE_f, visitno, cytokine, antigen) %>% 
    group_modify(.f = ~ broom::tidy(lm(value ~ EventIndPrimary1, data = .x), conf.int = TRUE)),
  # to get a tidy dataframe of model summaries
  y = ds.tmp.ics.tbl1 %>%
    group_by(VACCINE_f, visitno, cytokine, antigen) %>%
    group_modify(.f = ~ broom::glance(lm(value ~ EventIndPrimary1, data = .x))),
  by = c("VACCINE_f", "visitno", "cytokine", "antigen")
) %>%
  dplyr::ungroup(x = .) %>%
  filter(term == "EventIndPrimary1Non-cases") %>%
  mutate(`GM Ratio` = round(10^estimate, 2),
         `GM Ratio CI` = paste0("(", round(10^conf.low, 2), ", ", round(10^conf.high, 2), ")")) %>%
  select(VACCINE_f, visitno, cytokine, antigen, `GM Ratio`, `GM Ratio CI`) 



resp.diff.ci <- ds.tmp.ics.tbl1 %>%
  group_by(visitno, cytokine, antigen , EventIndPrimary1, VACCINE_f) %>%
  summarize(n = n(),
            r = sum(response))%>% 
  pivot_wider(names_from = EventIndPrimary1, values_from = c( n, r))

  resp.ci <- data_frame()

  for(i in 1:nrow(resp.diff.ci)){
    
    ds.tmp <-resp.diff.ci[i,] 
    
    
    resp.ci.tmp <- ds.tmp %>%
  mutate(lcl = prop.test(x = c(ds.tmp$r_Cases, ds.tmp$`r_Non-cases`), 
                               n = c(ds.tmp$n_Cases, ds.tmp$`n_Non-cases`), 
                               alternative = "two.sided", 
                               correct = FALSE)$conf.int[1],
         
         ucl = prop.test(x = c(ds.tmp$r_Cases, ds.tmp$`r_Non-cases`), 
                               n = c(ds.tmp$n_Cases, ds.tmp$`n_Non-cases`), 
                               alternative = "two.sided", 
                               correct = FALSE)$conf.int[2]) 
    
    resp.ci <- rbind(resp.ci, resp.ci.tmp)
    
  }
  
  
  resp.ci <- resp.ci %>%
    mutate(lcl =round(lcl*100,2),
           ucl =round(ucl*100, 2)) %>%
    mutate(`resp CI` = paste0("(", lcl, "%", ", ", ucl, "%)")) %>%
    select(visitno,cytokine, antigen, VACCINE_f, `resp CI`)



tbl2 <- full_join(tbl2.1, tbl2.2) %>%
  full_join(., resp.ci) %>%
  full_join(., tbl2.3) %>%
  mutate(VACCINE_f = factor(VACCINE_f, levels = c("Hybrid", "Vaccine")),
         viaitno = factor(visitno, levels = c("M0", "M1")),
         antigen = factor(antigen, levels = c("BA.4-5 N", "Spike BA.4/5", "Spike Wuhan"))) %>%
  arrange(VACCINE_f, viaitno, antigen) %>%
  select(VACCINE_f, viaitno,cytokine, antigen, n_Cases, `Response Rate_Cases`, `95% CI_Cases`, `Geometric Mean (GM)_Cases`, `GM 95% CI_Cases`, `n_Non-cases`, `Response Rate_Non-cases`, `95% CI_Non-cases`,
         `Geometric Mean (GM)_Non-cases`, `GM 95% CI_Non-cases`,
          `resp.diff`, `resp CI`, `GM Ratio`, `GM Ratio CI`)
  



## report table in RTF format
rtf.w <- 11
 rtf.h <- 8.5
 rtf.fs <- 11
 rtf.mar <- c(1,1,1,1)
 rtf.na <- '0' 
 out.rtf <- "../report/tbl1_ics.doc"
 rtf <- RTF(out.rtf, 
            width = rtf.w, height = rtf.h, 
            font.size = rtf.fs, omi = rtf.mar)
 tbl_lab <- paste0("Table 1. Immune marker response frequencies at M0 and vaccine-response frequencies at Peak (when applicable) by COVID-19 outcome status, geometric means (GMs) by COVID-19 outcome status, and their comparisons between non-cases and cases ")
 
 rtf.tbl <- tbl2 
 
  addHeader(rtf, tbl_lab,
          font.size = 12)
addTable(rtf, rtf.tbl,col.justify = c("L","L","L","C","C","C","C","C","C","C","C","C","C","C","C","C", "C","C"))
addPageBreak(rtf, 
             width = rtf.w, height = rtf.h, 
             font.size = rtf.fs, omi = rtf.mar)
rtf::done(rtf)






















