---
title: "- ICS - CD4+ T - High-Dimensional ICS Analysis -"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(cowplot)
library(data.table)
library(doMC)
```

# - Clustering
______________

Fluorescence intensity expression from CD4+ T expressing 1 of 8 cytokines or functional activation markers (IFN-$\gamma$, TNF, CD154, IL-2, IL-4, IL-5/IL-13, IL-17a and IL-21) were extracted; and Leiden clustering was performed.  

The gate used is the following:  

* CD4+ T - `/Time/K1/K2/K3/K4/K5/K6/K7/K8/Lv/14-/L/S/19-/3+/3+excl 16br/56-16-/4+`  

Duplicates, unreliable as well as out of window samples were removed.  

**Immunogenicity Set**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Data - CD4+ T - Immunogenicity Set
##-- meta.data
#- Unreliable PTIDs
unreliable.s <- readxl::read_xlsx(path = "./CoVPN-3008_Immunogenicity_Unreliable-data_18SEP23_VV.xlsx")
unreliable.s <- unreliable.s[2:nrow(unreliable.s), ]
unreliable.s %>%
  filter(`Unreliable Stim` == "ALL") %>%
  mutate(SAMPLE = paste(Batch, PTID, VISITNO, `Run Num`, sep = "_")) %>%
  pull(SAMPLE) -> unreliable.s_1
unreliable.s %>%
  filter(`Unreliable Stim` == "negctrl") %>%
  mutate(SAMPLE = paste(Batch, PTID, VISITNO, `Run Num`, Replicate, "negctrl", sep = "_")) %>%
  pull(SAMPLE) -> unreliable.s_2
unreliable.s %>%
  filter(!(`Unreliable Stim` %in% c("ALL", "negctrl"))) %>%
  mutate(SAMPLE = paste(Batch, PTID, VISITNO, `Run Num`, "1", `Unreliable Stim`, sep = "_")) %>%
  pull(SAMPLE) -> unreliable.s_3
#- Filtered PTIDs (Immunogenicity flags x3)
meta.data <- read_csv(file = "./CoVPN-3008_ICS_group.csv")
meta.data %>%
  filter(immfl == "Y" & immincfl == "Y" & pprem6ifl == "Y") %>%
  pull(subject) -> filtered.s
#- Out of windows
out.of.wind <- read_csv(file = "./CoVPN-3008_ICS_PT_filtered_sample_20231228.csv")
out.of.wind %>%
  filter(reason == "Out of window") %>%
  mutate(sample = paste(ptid, visitno)) %>%
  pull(sample) %>%
  unique() -> filtered.s_2


##-- Data (ICS)
#- MTL
dt.MTL <- readxl::read_xlsx(path = "./3008 Immunogenicity MTL_B.xlsx") %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO))
colnames(dt.MTL)[1] <- "BATCH"
colnames(dt.MTL)[10] <- "RUNNUM"
#- dt
list.raw <- readRDS(file = "./001_list-dt_CD4+T_1-of-8-cytokines(1)_immunogenicity.rds")
lapply(X = list.raw, function(x) x$exprs) %>%
  bind_rows() %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO)) -> dt.exprs.raw
#- Filtering (rows)
ICSR::do_filtering(dt = dt.exprs.raw,
                   remove_dup = TRUE,
                   dt.MTL_dup = dt.MTL,
                   remove_participants = TRUE,
                   participants_list = "CTAC0036",
                   remove_nsub = FALSE,
                   cutoff_nsub = NULL,
                   flag_unreliable = TRUE,
                   unreliable_list_all = unreliable.s_1,
                   unreliable_list_by_stim = c(unreliable.s_2, unreliable.s_3)) -> dt.immunogenicity
dt.immunogenicity <- dt.immunogenicity %>%
  # filter(PTID %in% filtered.s) %>%
  mutate(reliable_flag = case_when(paste(PTID, VISITNO) %in% filtered.s_2 ~ "0",
                                   .default = reliable_flag))
dt.immunogenicity$reliable_flag %>%
  table() # 1,159,091 CD4+ T / 245 PTIDs (when filter(PTID %in% filtered.s) == it works!)
dt.immunogenicity <- dt.immunogenicity %>%
  filter(reliable_flag == "1") %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITNO, from = c("2", "3", "4"), to = c("M0", "M1", "M2")))
#- Filtering (cols)
cols <- c("FCS", "BATCH", "PTID", "STIM", "VISITNO", "VISITMONTH", "RUNNUM", "REPLICATE", "SAMP_ORD", "NSUB", "CYTNUM",
          colnames(dt.immunogenicity)[str_detect(string = colnames(dt.immunogenicity), pattern = "Time")],
          colnames(dt.immunogenicity)[str_detect(string = colnames(dt.immunogenicity), pattern = "asinh")])
# UViD, CD14, CD19, CD3, CD56, CD16, CD8 & CD4 are removed (/Time/K1/K2/K3/K4/K5/K6/K7/K8/Lv/14-/L/S/19-/3+/3+excl 16br/56-16-/4+)
# CXCR5 is removed (unreliable marker)
cols <- cols[-(c(33, 42, 44, 45, 49, 52, 54, 55, 57, 58, 69, 71, 72, 76, 79, 81, 82, 84, 85))]
dt.immunogenicity <- dt.immunogenicity %>%
  select(cols)


##-- Data summary (ICS)
#- dt
lapply(X = list.raw, function(x) x$cytnum) %>%
  bind_rows() %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO)) -> dt.cytnum.raw
#- Filtering (rows)
ICSR::do_filtering(dt = dt.cytnum.raw,
                   remove_dup = TRUE,
                   dt.MTL_dup = dt.MTL,
                   remove_participants = TRUE,
                   participants_list = "CTAC0036",
                   remove_nsub = FALSE,
                   cutoff_nsub = NULL,
                   flag_unreliable = FALSE) -> dt.summary.immunogenicity
dt.summary.immunogenicity <- dt.summary.immunogenicity %>%
  filter(!(paste(BATCH, PTID, VISITNO, RUNNUM, sep = "_") %in% unreliable.s_1)) %>%
  filter(!(paste(BATCH, PTID, VISITNO, RUNNUM, REPLICATE, STIM, sep = "_") %in% unreliable.s_2)) %>%
  filter(!(paste(BATCH, PTID, VISITNO, RUNNUM, REPLICATE, STIM, sep = "_") %in% unreliable.s_3)) %>%
  # filter((PTID %in% filtered.s)) %>%
  filter(!(paste(PTID, VISITNO) %in% filtered.s_2)) %>%
  filter(boolean_CYTNUM == TRUE) %>%
  group_by(BATCH, PTID, STIM, VISITNO, RUNNUM, SAMP_ORD) %>%
  summarise(NSUB = sum(NSUB), CYTNUM = sum(CYTNUM)) %>%
  ungroup() %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITNO, from = c("2", "3", "4"), to = c("M0", "M1", "M2"))) # 245 PTIDs
```   
  
**Case cohort Set**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Data - CD4+ T - Case cohort Set
##-- meta.data
#- Unreliable PTIDs
unreliable.s <- readxl::read_xlsx(path = "./CoVPN 3008 Case Cohort_Unreliable data_16NOV23_A_modified_08FEB2024.xlsx")
unreliable.s <- unreliable.s[2:nrow(unreliable.s), ]
unreliable.s %>%
  filter(`Unreliable Stim` == "ALL") %>%
  mutate(SAMPLE = paste(`Batch #`, PTID, VISITNO, `Run Num`, sep = "_")) %>%
  pull(SAMPLE) -> unreliable.s_1
unreliable.s %>%
  filter(!(`Unreliable Stim` %in% c("ALL", "negctrl"))) %>%
  mutate(SAMPLE = paste(`Batch #`, PTID, VISITNO, `Run Num`, "1", `Unreliable Stim`, sep = "_")) %>%
  pull(SAMPLE) -> unreliable.s_2


##-- Data (ICS)
#- MTL
dt.MTL <- readxl::read_xlsx(path = "./CoVPN3008 Cases_MTL_A_14Sep23_VV.xlsx") %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "3.4" ~ "3", 
                             VISITNO == "4.2" ~ "4", 
                             VISITNO == "4.5999999999999996" ~ "4", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO))
colnames(dt.MTL)[1] <- "BATCH"
colnames(dt.MTL)[10] <- "RUNNUM"
#- dt
list.raw <- readRDS(file = "./001_list-dt_CD4+T_1-of-8-cytokines(1).rds")
lapply(X = list.raw, function(x) x$exprs) %>%
  bind_rows() %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "3.4" ~ "3", 
                             VISITNO == "4.2" ~ "4", 
                             VISITNO == "4.5999999999999996" ~ "4", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO)) -> dt.exprs.raw
#- Filtering (rows)
ICSR::do_filtering(dt = dt.exprs.raw,
                   remove_dup = TRUE,
                   dt.MTL_dup = dt.MTL,
                   remove_participants = TRUE,
                   participants_list = "CTAC0036",
                   remove_nsub = FALSE,
                   cutoff_nsub = NULL,
                   flag_unreliable = TRUE,
                   unreliable_list_all = unreliable.s_1,
                   unreliable_list_by_stim = unreliable.s_2) -> dt.cc
dt.cc$reliable_flag %>%
  table() # 195,069 CD4+ T / 67 PTIDs
dt.cc <- dt.cc %>%
  filter(reliable_flag == "1") %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITNO, from = c("2", "3", "4"), to = c("M0", "M1", "M2")))
#- Filtering (cols)
cols <- c("FCS", "BATCH", "PTID", "STIM", "VISITNO", "VISITMONTH", "RUNNUM", "REPLICATE", "SAMP_ORD", "NSUB", "CYTNUM",
          colnames(dt.cc)[str_detect(string = colnames(dt.cc), pattern = "Time")],
          colnames(dt.cc)[str_detect(string = colnames(dt.cc), pattern = "asinh")])
# UViD, CD14, CD19, CD3, CD56, CD16, CD8 & CD4 are removed (/Time/K1/K2/K3/K4/K5/K6/K7/K8/Lv/14-/L/S/19-/3+/3+excl 16br/56-16-/4+)
# CXCR5 is removed (unreliable marker)
cols <- cols[-(c(33, 42, 44, 45, 49, 52, 54, 55, 57, 58, 69, 71, 72, 76, 79, 81, 82, 84, 85))]
dt.cc <- dt.cc %>%
  select(cols)

##-- Data summary (ICS)
#- dt
lapply(X = list.raw, function(x) x$cytnum) %>%
  bind_rows() %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "3.4" ~ "3", 
                             VISITNO == "4.2" ~ "4", 
                             VISITNO == "4.5999999999999996" ~ "4", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO)) -> dt.cytnum.raw
#- Filtering (rows)
ICSR::do_filtering(dt = dt.cytnum.raw,
                   remove_dup = TRUE,
                   dt.MTL_dup = dt.MTL,
                   remove_participants = TRUE,
                   participants_list = "CTAC0036",
                   remove_nsub = FALSE,
                   cutoff_nsub = NULL,
                   flag_unreliable = TRUE,
                   unreliable_list_all = unreliable.s_1,
                   unreliable_list_by_stim = unreliable.s_2) -> dt.summary.cc
dt.summary.cc$reliable_flag %>%
  table() # 67 PTIDs
dt.summary.cc <- dt.summary.cc %>%
  filter(reliable_flag == "1") %>%
  filter(boolean_CYTNUM == TRUE) %>%
  filter(!is.na(PTID)) %>%
  group_by(BATCH, PTID, STIM, VISITNO, RUNNUM, SAMP_ORD) %>%
  summarise(NSUB = sum(NSUB), CYTNUM = sum(CYTNUM)) %>%
  ungroup() %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITNO, from = c("2", "3", "4"), to = c("M0", "M1", "M2"))) # 245 PTIDs
```   

**Merging - Immunogenicity x Case Cohort**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Merging
#- Merge
identical(colnames(dt.immunogenicity), colnames(dt.cc)) # TRUE
dt <- bind_rows(dt.immunogenicity %>%
                  mutate(SET = "Immunogenicity set"),
                dt.cc %>%
                  mutate(SET = "Case cohort set"))
identical(colnames(dt.summary.immunogenicity), colnames(dt.summary.cc)) # TRUE
dt.summary_Rhino <- bind_rows(dt.summary.immunogenicity %>%
                                mutate(SET = "Immunogenicity set"),
                              dt.summary.cc %>%
                                mutate(SET = "Case cohort set"))
intersect(unique(dt.immunogenicity$PTID), unique(dt.cc$PTID))

#- Characteristics
dt.nonfilt <- dt
dt.nonfilt <- dt.nonfilt %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2, 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))
table(dt.nonfilt$RX_CODE, dt.nonfilt$COHORT, useNA = "always")
dt.nonfilt %>%
  filter(SET == "Case cohort set") %>%
  filter(!duplicated(PTID)) %>%
  pull(COHORT) %>%
  table() # 2 non-cases; 65 cases
dt.nonfilt %>%
  filter(SET == "Case cohort set") %>%
  filter(!duplicated(PTID)) %>%
  pull(PTID) %>%
  unique() %>%
  length() # 67
dt.nonfilt %>%
  filter(SET != "Case cohort set") %>%
  filter(!duplicated(PTID)) %>%
  pull(COHORT) %>%
  table() # 264 non-cases; 5 cases
dt.nonfilt %>%
  filter(SET != "Case cohort set") %>%
  filter(!duplicated(PTID)) %>%
  pull(PTID) %>%
  unique() %>%
  length() # 278

#- Characteristics - check wth MTL
dt.MTL.imm <- readxl::read_xlsx(path = "./3008 Immunogenicity MTL_B.xlsx") %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2, 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))
dt.MTL.imm %>%
  filter(!duplicated(PTID)) %>%
  pull(COHORT) %>%
  table() # 286 non-cases; 6 cases
length(unique(dt.MTL.imm$PTID)) # 302
dt.MTL.cc <- readxl::read_xlsx(path = "./CoVPN3008 Cases_MTL_A_14Sep23_VV.xlsx") %>% 
  mutate(VISITNO = case_when(VISITNO == "3.1" ~ "3", 
                             VISITNO == "10.1" ~ "10", 
                             .default = VISITNO)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2, 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))
dt.MTL.cc %>%
  filter(!duplicated(PTID)) %>%
  pull(COHORT) %>%
  table() # 87 cases; 2 non-cases
length(unique(dt.MTL.cc$PTID)) # 90

#- Case-cohort info
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
  filter(FASFL == "Y") # 168 PTIDs
table(meta.data$Perprotocol)
table(meta.data$AGRP2, meta.data$EventIndPrimary1)
# 1: 22 x 7 (VIPP HIV+) - Total 29
# 2.1: 39 x 22 (HIPP HIV+) - Total 61
# 3: 21 x 0 (VIPP HIV-) - Total 21
# 4.1: 56 x 1 (HIPP HIV-) - Total 57
meta.data %>%
  pull(Subjectid) -> PTIDs
PTIDs <- str_remove(string = PTIDs, pattern = "C3008[-]") # 168 participants
intersect(unique(dt$PTID), PTIDs) %>% length() # 168 participants

#- Venn diagram
x <- list("Immunogenicity set" = dt %>% filter(SET == "Immunogenicity set") %>% pull(PTID) %>% unique(), 
          "Case-cohort set" = PTIDs)
ggvenn::ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF"),
  stroke_size = 0.5, set_name_size = 4
  ) # pdf 5 x 6

#- Filtering
dt <- dt %>%
  filter(PTID %in% PTIDs)
length(unique(dt$PTID)) # 168 participants
dt.summary_Rhino <- dt.summary_Rhino %>%
  filter(PTID %in% PTIDs)
length(unique(dt.summary_Rhino$PTID)) # 168 participants

#- Check (n's)
dt %>%
  filter(!duplicated(PTID)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2, 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE)) -> dt.check
table(dt.check$SET, dt.check$COHORT, dt.check$RX_CODE) # flowchart - ok!
dt %>%
  filter(!duplicated(paste(PTID, VISITNO, STIM))) %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2, 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE)) -> dt.check
table(dt.check$VISITNO, dt.check$STIM, dt.check$RX_CODE) # flowchart - ok!

#- Check (unreliables)
# Immunogenicity
dt %>%
  filter(paste(BATCH, PTID, VISITNO, RUNNUM, sep = "_") %in% unreliable.s_1)
dt %>%
  filter(paste(BATCH, PTID, VISITNO, RUNNUM, REPLICATE, STIM, sep = "_") %in% unreliable.s_2)
dt %>%
  filter(paste(BATCH, PTID, VISITNO, RUNNUM, REPLICATE, STIM, sep = "_") %in% unreliable.s_3)
# Case-cohort
dt %>%
  filter(paste(BATCH, PTID, VISITNO, RUNNUM, sep = "_") %in% unreliable.s_1)
dt %>%
  filter(paste(BATCH, PTID, VISITNO, RUNNUM, REPLICATE, STIM, sep = "_") %in% unreliable.s_2)

#- Output
gdata::keep(dt, dt.summary_Rhino, sure = TRUE)
```

**Leiden - k = 30 - asinh+asym**

Leiden clustering was made on the cluster.

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Leiden
#- k = 30 & resolution = 0.8 (asinh+asym)
partition <- readRDS(file = "./leiden-partition_k=30_niter=100_res=0.8_asinh+asym.rds")
dt$asinh_asym_Leiden <- partition
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Silhouette
#- Data
dt.ggplot <- readRDS(file = "./dt_silhouette_k=30_niter=100.rds")[[8]]

#- Fig.
my.colors <- scales::hue_pal()(16)
names(my.colors) <- paste("Leiden:", 1:16)
dt.ggplot %>%
  mutate(Leiden = factor(x = cluster, levels = paste("Leiden:", 1:16))) %>%
  arrange(Leiden, width) %>%
  mutate(cell.id = 1:nrow(dt.ggplot)) %>%
  ggplot() +
    geom_bar(aes(x = cell.id, weight = width, fill = Leiden)) +
    scale_fill_manual(values = my.colors) +
    labs(x = NULL, y = "Silhouette width", fill = NULL) +
    theme_bw() +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          panel.grid = element_blank(),
          legend.position = "bottom") -> plot
plot
# Silhouette: allowing us to quickly identify poorly separate clusters with mostly negative widths (such as 12)
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Purity
#- Data
dt.ggplot <- readRDS(file = "./dt_purity_k=30_niter=100.rds")[[8]]

#- Fig.
my.colors <- scales::hue_pal()(16)
names(my.colors) <- paste("Leiden:", 1:16)
dt.ggplot %>%
  mutate(Leiden = factor(x = cluster, levels = paste("Leiden:", 1:16))) %>%
  arrange(Leiden, purity) %>%
  mutate(cell.id = 1:nrow(dt.ggplot)) %>%
  ggplot() +
    geom_bar(aes(x = cell.id, weight = purity, fill = Leiden)) +
    geom_hline(yintercept = 0.9, color = "grey50", lty = 2, lwd = .25) +
    scale_fill_manual(values = my.colors) +
    labs(x = NULL, y = "Purity", fill = NULL) +
    theme_bw() +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          panel.grid = element_blank(),
          legend.position = "bottom") -> plot
plot
# Median purity values are consistently greater than 0.9, indicating that most cells in each cluster are primarily surrounded by other cells from the same cluster. Some clusters have low purity values that may warrant more careful inspection - these probably represent closely related subpopulations.
```

**UMAP - asinh+asym**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
#- UMAP (asinh+asym)
markers <- colnames(dt)[str_detect(string = colnames(dt), pattern = "asinh_asym")]
set.seed(1234)
umap <- uwot::umap(X = dt[, ..markers], n_neighbors = 10, n_components = 2, min_dist = .1, verbose = TRUE)
dt$asinh_asym_UMAP_1 <- umap[, 1] %>% as.numeric()
dt$asinh_asym_UMAP_2 <- umap[, 2] %>% as.numeric()
```

**`colnames()`**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- data.table()
#- Ordering
cols.tmp <- colnames(dt)[str_detect(string = colnames(dt), pattern = "asinh")]
cols.tmp <- cols.tmp[!str_detect(string = cols.tmp, pattern = "UMAP|Leiden")]
cols <- c("FCS", "SET", "BATCH", "PTID", "STIM", "VISITNO", "VISITMONTH", "RUNNUM", "REPLICATE", "SAMP_ORD", "NSUB", "CYTNUM", 
          "asinh_asym_UMAP_1", "asinh_asym_UMAP_2", "asinh_asym_Leiden",
          colnames(dt)[str_detect(string = colnames(dt), pattern = "Time")],
          cols.tmp)
dt <- dt[, ..cols]
```


## - asinh+asym
_______________

### - UMAP
__________

**UMAP - BATCH**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
dt %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    # ggpointdensity::geom_pointdensity(size = .1, method = "kde2d") +
    geom_density_2d_filled() +
    geom_density_2d(colour = "black", linewidth = .1) +
    facet_wrap(~ paste("BATCH:", BATCH), ncol = 6) +
    scale_fill_viridis_d(option = "B") + # scale_fill_viridis_d(option = "E")
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          strip.text = element_text(size = 16),
          panel.grid = element_blank(),
          legend.position = "none") -> plot
png("UMAP-by-BATCH.png", width = 14, height = 15, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - PTID**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
dt %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_density_2d_filled() +
    geom_density_2d(colour = "black", linewidth = .1) +
    facet_wrap(~ paste("PTID:", PTID), ncol = 15) +
    scale_fill_viridis_d(option = "B") + 
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          strip.text = element_text(size = 16),
          panel.grid = element_blank(),
          legend.position = "none") -> plot
png("UMAP-by-PTID.png", width = 35, height = 32, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - GROUP**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
   filter(FASFL == "Y") # 168 PTIDs
dt %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2N, 
                                   warn_missing = FALSE)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = RX_CODE, 
                                   from = c("1", "2.1", "3", "4.1"), 
                                   to = c("AG1", "AG2.1", "AG3", "AG4.1"), 
                                   warn_missing = FALSE)) %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_density_2d_filled() +
    geom_density_2d(colour = "black", linewidth = .1) +
    facet_wrap(~ RX_CODE, ncol = 4) +
    scale_fill_viridis_d(option = "B") + 
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          strip.text = element_text(size = 16),
          panel.grid = element_blank(),
          legend.position = "none") -> plot
png("UMAP-by-GROUP.png", width = 15, height = 4.5, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - COHORT**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
   filter(FASFL == "Y") # 168 PTIDs
dt %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))  %>%
  mutate(COHORT = plyr::mapvalues(x = COHORT, 
                                  from = c("1", "0"), 
                                  to = c("case", "non-case"), 
                                  warn_missing = FALSE)) %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_density_2d_filled() +
    geom_density_2d(colour = "black", linewidth = .1) +
    facet_wrap(~ COHORT, ncol = 4) +
    scale_fill_viridis_d(option = "B") + 
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          strip.text = element_text(size = 16),
          panel.grid = element_blank(),
          legend.position = "none") -> plot
png("UMAP-by-COHORT.png", width = 7.5, height = 4.5, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - VISITMONTH**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
dt %>%
  mutate(VISITMONTH = factor(x = VISITMONTH, levels = c("M0", "M1", "M2"))) %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_density_2d_filled() +
    geom_density_2d(colour = "black", linewidth = .1) +
    facet_wrap(~ VISITMONTH, ncol = 3) +
    scale_fill_viridis_d(option = "B") +
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          panel.grid = element_blank(),
          strip.text = element_text(size = 16),
          legend.position = "none") -> plot
png("UMAP-by-VISITMONTH.png", width = 9, height = 3.4, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - ANTIGEN**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
dt %>%
  mutate(STIM = factor(x = STIM, levels = c("BA.1-2-4-5 E / BA.1-2-4-5 M", "BA.4-5 N", "BA.4-5 S1", "BA.4-5 S2", "COV2 CON S1", "COV2 CON S2", "Wuhan E Wuhan M", "Wuhan N", "negctrl"))) %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_density_2d_filled() +
    geom_density_2d(colour = "black", linewidth = .1) +
    facet_wrap(~ STIM, ncol = 3) +
    scale_fill_viridis_d(option = "B") +
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          panel.grid = element_blank(),
          strip.text = element_text(size = 16),
          legend.position = "none") -> plot
png("UMAP-by-ANTIGEN.png", width = 9, height = 9, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - CYTNUM**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
dt %>%
  arrange(CYTNUM) %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_point(aes(color = CYTNUM %>% as.character()), size = .1) +
    scale_color_manual(values = colorRamps::matlab.like2(n = 8)) +
    labs(x = "UMAP-1", y = "UMAP-2", color = "CYTNUM") +
    guides(colour = guide_legend(override.aes = list(size = 4, alpha = 1))) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 16),
          panel.grid = element_blank(),
          strip.text = element_text(size = 16),
          legend.text = element_text(size = 14),
          legend.position = "bottom") -> plot
png("UMAP-by-CYTNUM.png", width = 4.25, height = 5, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off()
```

**UMAP - Marker gates**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
#- Data
cols <- c("asinh_asym_UMAP_1", "asinh_asym_UMAP_2", colnames(dt)[str_detect(string = colnames(dt), pattern = "Time")])
dt.ggplot <- dt[, ..cols] %>%
  reshape::melt(id.vars = c("asinh_asym_UMAP_1", "asinh_asym_UMAP_2")) %>%
  mutate(variable = factor(x = variable, levels = c(colnames(dt)[str_detect(string = colnames(dt), pattern = "Time")])))

#- Figures
dt.ggplot %>%
  filter(value == TRUE) %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_point(size = .1, alpha = .2, color = "grey70", data = dt.ggplot %>% filter(value == FALSE)) +
    ggpointdensity::geom_pointdensity(size = .1, method = "kde2d") +
    facet_wrap(~ variable, ncol = 5) +
    viridis::scale_color_viridis() +
    labs(x = "UMAP-1", y = "UMAP-2") +
    theme_bw() +
    theme(legend.position = "none", 
          axis.text = element_blank(), 
          axis.ticks = element_blank(), 
          axis.title = element_text(size = 16), 
          panel.grid = element_blank(), 
          strip.text = element_text(size = 6)) -> plot
png("UMAP-by-marker-gates.png", width = 20, height = 20, units = "in", pointsize = 12, res = 300)
print(plot)
dev.off() 
```

**UMAP - Markers**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
markers <- colnames(dt)[str_detect(string = colnames(dt), pattern = "asinh_asym")]
markers <- markers[!str_detect(string = markers, pattern = "cyCombine")]
markers <- markers[-c(1:3)]
for(j in markers){
  # data.table()
  dt.ggplot <- dt[, c("asinh_asym_UMAP_1", "asinh_asym_UMAP_2", ..j)]
  colnames(dt.ggplot)[3] <- "value"
  dt.ggplot$marker <- plyr::mapvalues(x = j, 
                                      from = markers, 
                                      to = c("Perforin", "IL-5 or IL-13", "Ki67", "IL-4", "IL-2", "Granzyme B", "IFN-\u03B3", "CCR7", "CD25", "PD-1", "CCR6", "TNF", "CD45RA", "CD154", "IL-21", "CXCR3", "FoxP3", "IL-17a"),
                                      warn_missing = FALSE)
  
  # Figure
  dt.ggplot %>%
    ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
      geom_point(aes(color = value), size = .1) +
      facet_wrap(~ marker) +
      viridis::scale_color_viridis(option = "D", limits = c(-1, 7.5)) +
      labs(x = "UMAP-1", y = "UMAP-2", color = "arcsinh(x/500) value") +
      theme_bw() +
      theme(legend.position = "bottom", 
            axis.text = element_blank(), 
            axis.ticks = element_blank(), 
            axis.title = element_text(size = 16), 
            panel.grid = element_blank(), 
            strip.text = element_text(size = 16)) -> plot
  
  # Output
  png(paste0("UMAP-by-markers_", 
                  str_replace(string = str_remove_all(string = j, pattern = "asinh_asym_"), pattern = "[/]", replacement = "-or-"), 
                  ".png"),
      width = 6, height = 6.5, units = "in", pointsize = 12, res = 300)
  print(plot)
  dev.off()
}
```

**UMAP - `Leiden`**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- UMAP
my.colors <- scales::hue_pal()(16)
names(my.colors) <- paste("Leiden:", 1:16)

dt %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_point(aes(color = asinh_asym_Leiden), size = .1) +
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL, title = NULL) +
    scale_color_manual(values = my.colors) +
    guides(colour = guide_legend(override.aes = list(size = 3, alpha = 1))) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 14),
          panel.grid = element_blank(),
          legend.text = element_text(size = 12),
          legend.position = "bottom") -> plot
png(filename = "UMAP-by-LEIDEN(1).png", 
    width = 6, height = 7.5, units = "in", pointsize = 12, res = 300)
plot
dev.off()

dt %>%
  ggplot(aes(x = asinh_asym_UMAP_1, y = asinh_asym_UMAP_2)) +
    geom_point(aes(color = asinh_asym_Leiden), size = .1) +
    labs(x = "UMAP-1", y = "UMAP-2", color = NULL, title = NULL) +
    scale_color_manual(values = my.colors) +
    facet_wrap(~ asinh_asym_Leiden) +
    theme_bw() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(size = 14),
          panel.grid = element_blank(),
          strip.text = element_text(size = 16),
          legend.text = element_text(size = 12),
          legend.position = "none") -> plot
png(filename = "UMAP-by-LEIDEN(2).png", 
    width = 16, height = 16.5, units = "in", pointsize = 12, res = 300)
plot
dev.off()
```

### - Annotation
________________

**Dot-plot**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Dot-plot
##-- Data (dt.tmp)
from <- colnames(dt)[str_detect(string = colnames(dt), pattern = "Time")][-1]
to <- c("CCR7+", "CD45RA+", "Naive+", "CM+", "EM+", "TEMRA+",  "IFN-\u03B3+", "IL-2+", "IL-4+", "IL-5_IL-13+", "IL-17a+", "IL-21+", "CD154+", "TNF+", "Granzyme B+", "Perforin+", "CD25+", "PD-1+", "CCR6+", "CXCR3+", "Ki67+", "FoxP3+")
dt.tmp <- dt
colnames(dt.tmp)[str_detect(string = colnames(dt.tmp), pattern = "Time")][-1] <- to


##-- dt.ggplot (boolean)
dt.ggplot.boolean <- ICSR::dt.ggplot_boolean(dt = dt.tmp, flowjo_gates = to, cluster_col = "asinh_asym_Leiden")
dt.ggplot.boolean %>% View()


##-- Dendrogram
library(ggdendro)
set.seed(1234)
#- Data & dendrogram
correlation <- dt.ggplot.boolean %>%
  select(cluster, variable, marker_positivity) %>%
  reshape2::acast(cluster ~ variable, value.var = "marker_positivity") %>%
  t() %>%
  cor()
hclust.res <- hclust(as.dist(1 - correlation)) 
plot(as.dendrogram(hclust.res)) # c(11, 3, 8, 1, 4, 7, 2, 5, 16, 9, 13, 15, 10, 12, 6, 14)

#- Fig.
ddata <- dendro_data(hclust.res, type = "rectangle")
plot.dendrogram <- ggplot(segment(ddata)) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
  scale_x_continuous(expand = c(0, .6)) + 
  theme_bw() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_blank())


##-- Barplot
plot.barplot <- dt %>%
  group_by(asinh_asym_Leiden, .drop = FALSE) %>%
  summarize(n = n()) 
plot.barplot <- plot.barplot %>%
  mutate(Leiden = factor(x = asinh_asym_Leiden, 
                         levels = paste("Leiden:", c(11, 3, 8, 1, 4, 7, 2, 5, 16, 9, 13, 15, 10, 12, 6, 14)))) %>%
  ggplot(aes(x = Leiden, y = n, fill = Leiden)) +
    geom_bar(stat = "identity") +
    scale_y_continuous(trans = "sqrt") +
    scale_fill_manual(values = my.colors) +
    labs(y = "n()") +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(size = 18),
          axis.ticks.x = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 20),
          panel.grid = element_blank(),
          legend.position = "none")


##-- Boxplot
# Proportion of clusters within each BATCH (ex: in BATCH #1, Leiden 1 + 2 + 3 .. = 1)
dt.ggplot <- dt %>%
  group_by(BATCH) %>%
  mutate(n.total = n()) %>%
  group_by(BATCH, asinh_asym_Leiden, n.total, .drop = FALSE) %>%
  summarise(n = n()) %>%
  mutate(Freq = n / n.total)
# Adjusted proportion of BATCH within each cluster (ex: In Leiden 1, BATCH #1 + #2 + .. = 1) 
dt.ggplot <- dt.ggplot %>%
  group_by(asinh_asym_Leiden) %>%
  mutate(sum.Freq.PhenoGraph = sum(Freq, na.rm = TRUE)) %>%
  mutate(Freq_2 = Freq / sum.Freq.PhenoGraph) %>%
  mutate(Leiden = factor(x = asinh_asym_Leiden, 
                         levels = paste("Leiden:", c(11, 3, 8, 1, 4, 7, 2, 5, 16, 9, 13, 15, 10, 12, 6, 14)))) %>%
  arrange(BATCH, Leiden)
plot.boxplot <- dt.ggplot %>%
  ggplot(aes(x = Leiden, y = Freq_2)) +
    geom_boxplot(aes(fill = Leiden), outlier.shape = NA, alpha = .5) +
    geom_line(aes(group = BATCH), position = position_jitter(seed = 4, width = .2), color = "grey70", alpha = .5) +
    geom_point(aes(group = BATCH, fill = Leiden), 
               position = position_jitter(seed = 4, width = .2), 
               color = "grey60", size = 5, alpha = .8, pch = 21) +
    scale_fill_manual(values = my.colors) +
    scale_y_continuous(limits = c(0, 1), breaks = c(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
    labs(x = NULL, y = "Adjusted proportion of BATCH") +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.y = element_text(size = 18),
          axis.title = element_text(size = 20),
          panel.grid = element_blank(),
          legend.position = "none")


##-- Dotplot
#- flowjo_gates
flowjo_gates <- to

#- Markers 
from <- colnames(dt.tmp)[str_detect(string = colnames(dt.tmp), pattern = "asinh_asym")]
from <- from[!str_detect(string = from, pattern = "cyCombine")]
from <- from[-c(1:3)]
to <- c("asinh_asym_Perforin", "asinh_asym_IL-5_IL-13", "asinh_asym_Ki67", "asinh_asym_IL-4", "asinh_asym_IL-2", "asinh_asym_Granzyme B", "asinh_asym_IFN-\u03B3", "asinh_asym_CCR7", "asinh_asym_CD25", "asinh_asym_PD-1", "asinh_asym_CCR6", "asinh_asym_TNF", "asinh_asym_CD45RA", "asinh_asym_CD154", "asinh_asym_IL-21", "asinh_asym_CXCR3", "asinh_asym_FoxP3", "asinh_asym_IL-17a")
colnames(dt.tmp)[colnames(dt.tmp) %in% from] <- to

#- colnames subset & rename
cols <- c("asinh_asym_Leiden", 
          colnames(dt.tmp)[str_detect(string = colnames(dt.tmp), pattern = "[+]")][-1],
          to)
dt.tmp <- dt.tmp[, ..cols]
colnames(dt.tmp) <- str_remove(string = colnames(dt.tmp), pattern = "asinh_asym_")
markers <- colnames(dt.tmp)[24:41]

#- Dotplot
order_level_marker <- c("Naive", "CM", "EM", "TEMRA", "CCR7", "CD45RA", "IFN-γ", "IL-2", "IL-4", "IL-5_IL-13", "IL-17a", "IL-21", "CD154", "TNF", "Granzyme B", "Perforin", "CD25", "PD-1", "CCR6", "CXCR3", "Ki67", "FoxP3")
plot.dotplot <- ICSR::dotplot_annotation(
  dt = dt.tmp,
  markers = markers,
  flowjo_gates = flowjo_gates,
  cluster_col = "Leiden",
  level_marker = order_level_marker,
  order = paste("Leiden:", c(11, 3, 8, 1, 4, 7, 2, 5, 16, 9, 13, 15, 10, 12, 6, 14)))

#- annotation
annotation <- ICSR::x_axis(dt.boolean = dt.ggplot.boolean, 
                           order = paste("Leiden:", c(11, 3, 8, 1, 4, 7, 2, 5, 16, 9, 13, 15, 10, 12, 6, 14)))
annotation <- str_remove(string = annotation, pattern = "\nNA")

#- Output (.pdf) (1)
cairo_pdf(file = "dotplot-by-LEIDEN(1).pdf", width = 20, height = 25)
plot <- plot_grid(plot.dendrogram, 
                  plot.barplot, 
                  plot.dotplot + scale_x_discrete(labels = annotation),
                  align = "v", 
                  axis = "l",
                  ncol = 1, 
                  rel_heights = c(1, 1.5, 6))
print(plot)
dev.off()

#- Output (.pdf) (2)
cairo_pdf(file = "dotplot-by-LEIDEN(2).pdf", width = 20, height = 25)
plot <- plot_grid(plot.dendrogram, 
                  plot.boxplot, 
                  plot.dotplot + scale_x_discrete(labels = annotation),
                  align = "v", 
                  axis = "l",
                  ncol = 1, 
                  rel_heights = c(1, 1.5, 6))
print(plot)
dev.off()
```

**Ridge plot**

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Ridge plots
markers <- colnames(dt)[str_detect(string = colnames(dt), pattern = "asinh_asym")]
markers <- markers[!str_detect(string = markers, pattern = "cyCombine")]
markers <- markers[-c(1:3)]
cairo_pdf("ridge-plot-by-LEIDEN.pdf", width = 5, height = 10, onefile = TRUE)
for(marker in markers){
  # data.table()
  dt.ggplot <- dt[, ..marker] %>%
    mutate(marker = plyr::mapvalues(x = marker, 
                                    from = markers, 
                                    to = c("Perforin", "IL-5/IL-13", "Ki67", "IL-4", "IL-2", "GzB", "IFN-\u03B3", "CCR7", "CD25", "PD-1", "CCR6", "TNF", "CD45RA", "CD154", "IL-21", "CXCR3", "FOXP3", "IL-17A")))
  colnames(dt.ggplot)[1] <- "value"
  dt.ggplot$Leiden <- dt$asinh_asym_Leiden
  
  # Figure
  dt.ggplot %>%
    ggplot(aes(x = value, y = Leiden, fill = Leiden)) +
      ggridges::geom_density_ridges(bandwidth = .1, size = .2) +
      geom_vline(xintercept = 0, alpha = .5) +
      scale_x_continuous(limits = c(-1, 7.5)) +
      scale_fill_manual(values = my.colors) +
      facet_wrap(~ marker) +
      theme_bw() +
      labs(x = NULL, y = NULL) +
      theme(legend.position = "none", 
            axis.title = element_text(size = 16), 
            panel.grid = element_blank(), 
            strip.text = element_text(size = 16)) -> plot
  print(plot)
}
dev.off()
```

### - Positivity calls (`MIMOSA`)
_________________________________

For each cluster, STIM and VISITNO, a `MIMOSA` test (Finak et al., 2013) is applied to test whether the number of cytokine-producing cells for the stimulated data is equal to or greater than that for the negative control data. For each STIM, multiplicity adjustment is made.  
NB: We are blinded for the `RX_CODE`.  

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Data
#- Pre-processing
# Replicates are pooled together
dt.summary <- dt %>% 
  group_by(BATCH, PTID, STIM, VISITNO, RUNNUM, NSUB) %>%
  summarize(CYTNUM = n()) %>%
  group_by(BATCH, PTID, STIM, VISITNO, RUNNUM) %>%
  summarise(NSUB = sum(NSUB), CYTNUM = sum(CYTNUM)) %>%
  ungroup()

#- Samples without cytnum
setdiff(paste(dt.summary_Rhino$PTID, dt.summary_Rhino$STIM, dt.summary_Rhino$VISITNO),
        paste(dt.summary$PTID, dt.summary$STIM, dt.summary$VISITNO)) 

#- Pre-processing by Leiden
dt.tmp <- dt %>%
  group_by(BATCH, PTID, STIM, VISITNO, asinh_asym_Leiden, .drop = FALSE) %>%
  summarize(CYTNUM = n()) 
dt.tmp <- dt.tmp %>%
  mutate(NSUB = plyr::mapvalues(x = paste(BATCH, PTID, STIM, VISITNO), 
                                from = paste(dt.summary$BATCH, 
                                             dt.summary$PTID, 
                                             dt.summary$STIM, 
                                             dt.summary$VISITNO), 
                                to = dt.summary$NSUB,
                                warn_missing = FALSE)) %>%
  mutate(NSUB = NSUB %>% as.numeric()) %>%
  ungroup() %>%
  mutate(SAMPLE = paste(PTID, VISITNO, asinh_asym_Leiden)) %>%
  select(BATCH, PTID, STIM, VISITNO, SAMPLE, asinh_asym_Leiden, NSUB, CYTNUM)

#- Background subtraction
dt.tmp_1 <- dt.tmp %>%
  filter(STIM == "negctrl") %>%
  rename(NSUB_NEG = "NSUB", CYTNUM_NEG = "CYTNUM") %>%
  select(SAMPLE, NSUB_NEG, CYTNUM_NEG) 
dt.tmp_2 <- dt.tmp %>%
  filter(STIM != "negctrl")
dt.Leiden <- merge(x = dt.tmp_2, y = dt.tmp_1, by = "SAMPLE", all.x = TRUE) %>%
  mutate(PCTPOS = (CYTNUM / NSUB) * 100) %>%
  mutate(PCTNEG = (CYTNUM_NEG / NSUB_NEG) * 100) %>%
  mutate(PCTPOS_ADJ = PCTPOS - PCTNEG) %>%
  select(BATCH, PTID, STIM, VISITNO, asinh_asym_Leiden, NSUB, CYTNUM, PCTPOS, NSUB_NEG, CYTNUM_NEG, PCTNEG, PCTPOS_ADJ) %>%
  arrange(BATCH, PTID, STIM, VISITNO)
colnames(dt.Leiden)[5] <- "Leiden"
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Positivity calls
#- Data
mimosaSet <- dt.Leiden %>%
  mutate(SAMPLE = paste(PTID, VISITNO)) %>%
  select(PTID, STIM, VISITNO, SAMPLE, Leiden, NSUB, CYTNUM, NSUB_NEG, CYTNUM_NEG)
names(mimosaSet) <- toupper(names(mimosaSet))
write.table(x = mimosaSet, file = "in-mimosaSet.csv", row.names = FALSE, sep = ",")

#- Run MIMOSA
library(MIMOSA)
inFileName <- "in-mimosaSet.csv"
outFileName <- "out-mimosaSet.csv"
ICSR::runMIMOSA(INFILE = inFileName, OUTFILE = outFileName, CLUSTERS = paste("Leiden:", 1:16))

#- Processing
# For each STIM, multiplicity adjustment is made. == Adjustment is made by STIM.
dt.results_MIMOSA <- read_csv(file = "out-mimosaSet.csv")

#- Add positivity results
dt.Leiden <- dt.Leiden %>%
  mutate(MIMOSA_pr_nonresp = plyr::mapvalues(x = paste(PTID, VISITNO, STIM, Leiden), 
                                      from = paste(dt.results_MIMOSA$ptid, 
                                                   dt.results_MIMOSA$visitno, 
                                                   dt.results_MIMOSA$stim,
                                                   dt.results_MIMOSA$leiden),
                                      to = dt.results_MIMOSA$pr_nonresp, 
                                      warn_missing = FALSE)) %>%
  mutate(MIMOSA_pr_nonresp = MIMOSA_pr_nonresp %>% as.numeric()) %>%
  mutate(MIMOSA_pr_resp = plyr::mapvalues(x = paste(PTID, VISITNO, STIM, Leiden), 
                                      from = paste(dt.results_MIMOSA$ptid, 
                                                   dt.results_MIMOSA$visitno, 
                                                   dt.results_MIMOSA$stim,
                                                   dt.results_MIMOSA$leiden),
                                      to = dt.results_MIMOSA$pr_resp, 
                                      warn_missing = FALSE)) %>%
  mutate(MIMOSA_pr_resp = MIMOSA_pr_resp %>% as.numeric()) %>%
  mutate(MIMOSA_FDR = plyr::mapvalues(x = paste(PTID, VISITNO, STIM, Leiden), 
                                      from = paste(dt.results_MIMOSA$ptid, 
                                                   dt.results_MIMOSA$visitno, 
                                                   dt.results_MIMOSA$stim,
                                                   dt.results_MIMOSA$leiden),
                                      to = dt.results_MIMOSA$fdr, 
                                      warn_missing = FALSE)) %>%
  mutate(MIMOSA_FDR = MIMOSA_FDR %>% as.numeric()) %>%
  mutate(MIMOSA_call = ifelse(MIMOSA_pr_resp > 0.999, TRUE, FALSE))
dt.Leiden$MIMOSA_call %>% table(useNA = "always")
# dt.Leiden <- dt.Leiden %>%
#   mutate(MIMOSA_FDR = case_when(is.na(MIMOSA_call) == TRUE ~ NA, .default = MIMOSA_FDR))
```

### - Positivity calls (`fisher.test`)
______________________________________

For each cluster, STIM and VISITNO, a Fisher's exact test is applied to test whether the number of cytokine-producing cells for the stimulated data is equal to or greater than that for the negative control data. For each combination of PTID, VISITNO and LEIDEN, multiplicity adjustment is made over the number of ANTIGEN.  
NB: We are blinded for the `RX_CODE`.  

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Positivity calls (Fisher's exact test)
#- Data
dt.results_Fisher <- dt.Leiden %>% 
  mutate(SAMPLE = paste(PTID, VISITNO)) %>%
  select(PTID, STIM, VISITNO, SAMPLE, Leiden, NSUB, CYTNUM, NSUB_NEG, CYTNUM_NEG)
names(dt.results_Fisher) <- toupper(names(dt.results_Fisher))

#- Fisher's exact test
dt.results_Fisher$raw_p <- NA
for(i in 1:nrow(dt.results_Fisher))
{
  pval <- NA
  x.ant <- dt.results_Fisher[["CYTNUM"]][i]
  N.ant <- dt.results_Fisher[["NSUB"]][i] - dt.results_Fisher[["CYTNUM"]][i]
  x.neg <- dt.results_Fisher[["CYTNUM_NEG"]][i]
  N.neg <- dt.results_Fisher[["NSUB_NEG"]][i] - dt.results_Fisher[["CYTNUM_NEG"]][i]
  if(!is.na(N.neg) & N.ant > 0 & N.neg > 0) {
    m <- matrix(c(x.ant, N.ant, x.neg, N.neg), nc = 2, byrow = FALSE)
    pval <- fisher.test(m, alternative = "greater")$p.value
  }else{
    pval <- NA
  }
  dt.results_Fisher$raw_p[i] <- pval
}

#- Correction for multiple testing
dt.results_Fisher <- dt.results_Fisher %>%
  group_by(PTID, VISITNO, LEIDEN) %>%
  mutate(num_tests = n()) %>%
  mutate(adjvar = "ptid,visitno,leiden") %>%
  mutate(adjp = p.adjust(p = raw_p, method = "holm")) %>%
  ungroup()

#- Add positivity results
dt.Leiden <- dt.Leiden %>%
  mutate(FISHER_num.tests = plyr::mapvalues(x = paste(PTID, VISITNO, STIM, Leiden), 
                                            from = paste(dt.results_Fisher$PTID, 
                                                         dt.results_Fisher$VISITNO, 
                                                         dt.results_Fisher$STIM,
                                                         dt.results_Fisher$LEIDEN),
                                            to = dt.results_Fisher$num_tests, 
                                            warn_missing = FALSE)) %>%
  mutate(FISHER_adjvar = "ptid,visitno,leiden") %>%
  mutate(FISHER_FDR = plyr::mapvalues(x = paste(PTID, VISITNO, STIM, Leiden), 
                                      from = paste(dt.results_Fisher$PTID, 
                                                   dt.results_Fisher$VISITNO, 
                                                   dt.results_Fisher$STIM,
                                                   dt.results_Fisher$LEIDEN),
                                      to = dt.results_Fisher$adjp, 
                                      warn_missing = FALSE)) %>%
  mutate(FISHER_FDR = FISHER_FDR %>% as.numeric()) %>%
  mutate(FISHER_call = ifelse(FISHER_FDR < .00001, TRUE, FALSE))
dt.Leiden$FISHER_call %>% table(useNA = "always")

#- Comparison
table(dt.Leiden$MIMOSA_call, dt.Leiden$FISHER_call)
dt.Leiden %>%
  filter(MIMOSA_call == TRUE) %>%
  filter(FISHER_call == FALSE) %>%
  View()
```

### - Combined responses & Processing
_____________________________________

Combined peptide pool responses and magnitudes are calculated, including all pools regardless of positive or negative responder status. Negative magnitudes are censored at 0 prior to calculating the sum or maximum (across peptide pools); if the sum or maximum is < 0.001, the sum is set to 0.001 percent. Specifically, the combined overall magnitudes are defined as follows:  

* __Spike Ancestral__ = sum(COV2 CON S1, COV2 CON S2);  

* __Spike BA4/5__ = sum(BA.4-5 S1, BA.4-5 S2);  

* __Any Ancestral__ = sum(COV2 CON S1, COV2 CON S2, Wuhan N, Wuhan E/M);  

* __Any BA4/5__ = sum(BA.4-5 S1, BA.4-5 S2, BA.4-5 N, BA.1-2-4-5 E/M).    

If at least one peptide pool for a specific SARS-CoV-2 protein is positive (`MIMOSA` - FDR 1%), then the overall response to the protein is considered positive. After combined responses are calculated, for individual peptide pools magnitudes < 0.001 are set to 0.001.  

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Spike Ancestral
#- Sum of COV2 CON S1 and COV2 CON S2
dt.tmp <- dt.Leiden %>%
  filter(STIM %in% c("COV2 CON S1", "COV2 CON S2")) %>%
  group_by(PTID, VISITNO, Leiden) %>%
  filter(n() == 2) %>%
  mutate(PCTPOS_ADJ = ifelse(PCTPOS_ADJ < 0, 0, PCTPOS_ADJ)) %>%
  summarize(PCTPOS_ADJ = sum(PCTPOS_ADJ, na.rm = TRUE)) %>%
  mutate(STIM = "Spike Ancestral")
#- Add Status (MIMOSA)
samples <- dt.Leiden %>%
  filter(STIM %in% c("COV2 CON S1", "COV2 CON S2")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  mutate(SAMPLE = paste(PTID, VISITNO, Leiden)) %>%
  pull(SAMPLE) %>% 
  unique()
dt.tmp <- dt.tmp %>%
  mutate(MIMOSA_call = ifelse(paste(PTID, VISITNO, Leiden) %in% samples, TRUE, FALSE))
#- Add results to table
dt.tmp <- dt.tmp %>%
  mutate(BATCH = NA) %>%
  mutate(NSUB = NA) %>%
  mutate(CYTNUM = NA) %>%
  mutate(PCTPOS = NA) %>%
  mutate(NSUB_NEG = NA) %>%
  mutate(CYTNUM_NEG = NA) %>%
  mutate(PCTNEG = NA) %>%
  mutate(MIMOSA_pr_nonresp = NA) %>%
  mutate(MIMOSA_pr_resp = NA) %>%
  mutate(MIMOSA_FDR = NA) %>%
  mutate(FISHER_num.tests = NA) %>%
  mutate(FISHER_adjvar = NA) %>%
  mutate(FISHER_FDR = NA) %>%
  mutate(FISHER_call = NA) %>%
  select(BATCH, PTID, STIM, VISITNO, Leiden, NSUB, CYTNUM, PCTPOS, NSUB_NEG, CYTNUM_NEG, PCTNEG, PCTPOS_ADJ, MIMOSA_pr_nonresp, MIMOSA_pr_resp, MIMOSA_FDR, MIMOSA_call, FISHER_num.tests, FISHER_adjvar, FISHER_FDR, FISHER_call)
identical(colnames(dt.Leiden), colnames(dt.tmp)) # TRUE
dt.Leiden <- bind_rows(dt.Leiden, dt.tmp)
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Spike BA4/5
#- Sum of BA.4-5 S1 and BA.4-5 S2
dt.tmp <- dt.Leiden %>%
  filter(STIM %in% c("BA.4-5 S1", "BA.4-5 S2")) %>%
  group_by(PTID, VISITNO, Leiden) %>%
  filter(n() == 2) %>%
  mutate(PCTPOS_ADJ = ifelse(PCTPOS_ADJ < 0, 0, PCTPOS_ADJ)) %>%
  summarize(PCTPOS_ADJ = sum(PCTPOS_ADJ, na.rm = TRUE)) %>%
  mutate(STIM = "Spike BA4/5")
#- Add Status (MIMOSA)
samples <- dt.Leiden %>%
  filter(STIM %in% c("BA.4-5 S1", "BA.4-5 S2")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  mutate(SAMPLE = paste(PTID, VISITNO, Leiden)) %>%
  pull(SAMPLE) %>% 
  unique()
dt.tmp <- dt.tmp %>%
  mutate(MIMOSA_call = ifelse(paste(PTID, VISITNO, Leiden) %in% samples, TRUE, FALSE))
#- Add results to table
dt.tmp <- dt.tmp %>%
  mutate(BATCH = NA) %>%
  mutate(NSUB = NA) %>%
  mutate(CYTNUM = NA) %>%
  mutate(PCTPOS = NA) %>%
  mutate(NSUB_NEG = NA) %>%
  mutate(CYTNUM_NEG = NA) %>%
  mutate(PCTNEG = NA) %>%
  mutate(MIMOSA_pr_nonresp = NA) %>%
  mutate(MIMOSA_pr_resp = NA) %>%
  mutate(MIMOSA_FDR = NA) %>%
  mutate(FISHER_num.tests = NA) %>%
  mutate(FISHER_adjvar = NA) %>%
  mutate(FISHER_FDR = NA) %>%
  mutate(FISHER_call = NA) %>%
  select(BATCH, PTID, STIM, VISITNO, Leiden, NSUB, CYTNUM, PCTPOS, NSUB_NEG, CYTNUM_NEG, PCTNEG, PCTPOS_ADJ, MIMOSA_pr_nonresp, MIMOSA_pr_resp, MIMOSA_FDR, MIMOSA_call, FISHER_num.tests, FISHER_adjvar, FISHER_FDR, FISHER_call)
identical(colnames(dt.Leiden), colnames(dt.tmp)) # TRUE
dt.Leiden <- bind_rows(dt.Leiden, dt.tmp)
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Any Ancestral
#- Sum of COV2 CON S1, COV2 CON S2, Wuhan N and Wuhan E/M
dt.tmp <- dt.Leiden %>%
  filter(STIM %in% c("COV2 CON S1", "COV2 CON S2", "Wuhan N", "Wuhan E Wuhan M")) %>%
  group_by(PTID, VISITNO, Leiden) %>%
  filter(n() == 4) %>%
  mutate(PCTPOS_ADJ = ifelse(PCTPOS_ADJ < 0, 0, PCTPOS_ADJ)) %>%
  summarize(PCTPOS_ADJ = sum(PCTPOS_ADJ, na.rm = TRUE)) %>%
  mutate(STIM = "Any Ancestral")
#- Add Status (MIMOSA)
samples <- dt.Leiden %>%
  filter(STIM %in% c("COV2 CON S1", "COV2 CON S2", "Wuhan N", "Wuhan E Wuhan M")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  mutate(SAMPLE = paste(PTID, VISITNO, Leiden)) %>%
  pull(SAMPLE) %>% 
  unique()
dt.tmp <- dt.tmp %>%
  mutate(MIMOSA_call = ifelse(paste(PTID, VISITNO, Leiden) %in% samples, TRUE, FALSE))
#- Add results to table
dt.tmp <- dt.tmp %>%
  mutate(BATCH = NA) %>%
  mutate(NSUB = NA) %>%
  mutate(CYTNUM = NA) %>%
  mutate(PCTPOS = NA) %>%
  mutate(NSUB_NEG = NA) %>%
  mutate(CYTNUM_NEG = NA) %>%
  mutate(PCTNEG = NA) %>%
  mutate(MIMOSA_pr_nonresp = NA) %>%
  mutate(MIMOSA_pr_resp = NA) %>%
  mutate(MIMOSA_FDR = NA) %>%
  mutate(FISHER_num.tests = NA) %>%
  mutate(FISHER_adjvar = NA) %>%
  mutate(FISHER_FDR = NA) %>%
  mutate(FISHER_call = NA) %>%
  select(BATCH, PTID, STIM, VISITNO, Leiden, NSUB, CYTNUM, PCTPOS, NSUB_NEG, CYTNUM_NEG, PCTNEG, PCTPOS_ADJ, MIMOSA_pr_nonresp, MIMOSA_pr_resp, MIMOSA_FDR, MIMOSA_call, FISHER_num.tests, FISHER_adjvar, FISHER_FDR, FISHER_call)
identical(colnames(dt.Leiden), colnames(dt.tmp)) # TRUE
dt.Leiden <- bind_rows(dt.Leiden, dt.tmp)
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Any BA4/5
#- Sum of BA.4-5 S1, BA.4-5 S2, BA.4-5 S1, BA.4-5 N and BA.1-2-4-5 E/M
dt.tmp <- dt.Leiden %>%
  filter(STIM %in% c("BA.4-5 S1", "BA.4-5 S2", "BA.4-5 N", "BA.1-2-4-5 E / BA.1-2-4-5 M")) %>%
  group_by(PTID, VISITNO, Leiden) %>%
  filter(n() == 4) %>%
  mutate(PCTPOS_ADJ = ifelse(PCTPOS_ADJ < 0, 0, PCTPOS_ADJ)) %>%
  summarize(PCTPOS_ADJ = sum(PCTPOS_ADJ, na.rm = TRUE)) %>%
  mutate(STIM = "Any BA4/5")
#- Add Status (MIMOSA)
samples <- dt.Leiden %>%
  filter(STIM %in% c("BA.4-5 S1", "BA.4-5 S2", "BA.4-5 N", "BA.1-2-4-5 E / BA.1-2-4-5 M")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  mutate(SAMPLE = paste(PTID, VISITNO, Leiden)) %>%
  pull(SAMPLE) %>% 
  unique()
dt.tmp <- dt.tmp %>%
  mutate(MIMOSA_call = ifelse(paste(PTID, VISITNO, Leiden) %in% samples, TRUE, FALSE))
#- Add results to table
dt.tmp <- dt.tmp %>%
  mutate(BATCH = NA) %>%
  mutate(NSUB = NA) %>%
  mutate(CYTNUM = NA) %>%
  mutate(PCTPOS = NA) %>%
  mutate(NSUB_NEG = NA) %>%
  mutate(CYTNUM_NEG = NA) %>%
  mutate(PCTNEG = NA) %>%
  mutate(MIMOSA_pr_nonresp = NA) %>%
  mutate(MIMOSA_pr_resp = NA) %>%
  mutate(MIMOSA_FDR = NA) %>%
  mutate(FISHER_num.tests = NA) %>%
  mutate(FISHER_adjvar = NA) %>%
  mutate(FISHER_FDR = NA) %>%
  mutate(FISHER_call = NA) %>%
  select(BATCH, PTID, STIM, VISITNO, Leiden, NSUB, CYTNUM, PCTPOS, NSUB_NEG, CYTNUM_NEG, PCTNEG, PCTPOS_ADJ, MIMOSA_pr_nonresp, MIMOSA_pr_resp, MIMOSA_FDR, MIMOSA_call, FISHER_num.tests, FISHER_adjvar, FISHER_FDR, FISHER_call)
identical(colnames(dt.Leiden), colnames(dt.tmp)) # TRUE
dt.Leiden <- bind_rows(dt.Leiden, dt.tmp)
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Processing
#- Arrange
dt.Leiden <- dt.Leiden %>%
  mutate(PCTPOS_ADJ_noadjustment = PCTPOS_ADJ) %>%
  mutate(PCTPOS_ADJ = ifelse(PCTPOS_ADJ < 0.001, 0.001, PCTPOS_ADJ)) %>%
  mutate(Leiden = factor(x = Leiden, levels = paste("Leiden:", 1:16))) %>%
  mutate(STIM = factor(x = STIM, levels = c("Any Ancestral", "Any BA4/5", "Spike Ancestral", "Spike BA4/5", "COV2 CON S1", "COV2 CON S2", "Wuhan N", "Wuhan E Wuhan M", "BA.4-5 S1", "BA.4-5 S2", "BA.4-5 N", "BA.1-2-4-5 E / BA.1-2-4-5 M"))) %>%
  arrange(PTID, STIM, VISITNO, Leiden) %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITNO, from = c("2", "3", "4"), to = c("M0", "M1", "M2")))

#- Ordering
cols <- c("BATCH", "PTID", "STIM", "VISITNO", "VISITMONTH", "Leiden", 
          "NSUB", "CYTNUM", "PCTPOS", 
          "NSUB_NEG", "CYTNUM_NEG", "PCTNEG", 
          "PCTPOS_ADJ_noadjustment", "PCTPOS_ADJ", 
          "MIMOSA_pr_nonresp", "MIMOSA_pr_resp", "MIMOSA_FDR", "MIMOSA_call",
          "FISHER_num.tests", "FISHER_adjvar", "FISHER_FDR", "FISHER_call")
dt.Leiden <- dt.Leiden[, cols]

#- Output (.csv)
dt.Leiden %>%
  write_csv(file = "./dt_Leiden.csv")
```

### - Box-plots
_______________

**Exploratory Data Analysis**

All clusters are evaluated in an exploratory data analysis to ensure that the importance of rare cell subsets is not overlooked.  

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Box-plots
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
   filter(FASFL == "Y") # 168 PTIDs
dt.Leiden <- dt.Leiden %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2N, 
                                   warn_missing = FALSE)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = RX_CODE, 
                                   from = c("1", "2.1", "3", "4.1"), 
                                   to = c("AG1", "AG2-1", "AG3", "AG4-1"), 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))  %>%
  mutate(COHORT = plyr::mapvalues(x = COHORT, 
                                  from = c("1", "0"), 
                                  to = c("case", "non-case"), 
                                  warn_missing = FALSE))

#- dt.ggplot
dt.ggplot <- dt.Leiden %>%
  ungroup() %>%
  filter(!(is.na(MIMOSA_call))) %>%
  mutate(VISITMONTH = factor(x = VISITMONTH, levels = c("M0", "M1", "M2"))) %>%
  mutate(FILL_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "TRUE" ~ "#A6CEE3",
                              RX_CODE %in% "AG2-1" & MIMOSA_call == "TRUE" ~ "#B2DF8A",
                              RX_CODE %in% "AG3" & MIMOSA_call == "TRUE" ~ "#FB9A99",
                              RX_CODE %in% "AG4-1" & MIMOSA_call == "TRUE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "FALSE" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" & MIMOSA_call == "FALSE" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" & MIMOSA_call == "FALSE" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" & MIMOSA_call == "FALSE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_LINE = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                                RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                                RX_CODE %in% "AG3" ~ "#FB9A99",
                                RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                                .default = "white")) %>%
  mutate(FILL_BX = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                             RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                             RX_CODE %in% "AG3" ~ "#FB9A99",
                             RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                             .default = "white")) %>%
  mutate(FILL_BX = factor(x = FILL_BX, levels = c("#FB9A99", "#FDBF6F", "#A6CEE3", "#B2DF8A"))) %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITMONTH, from = c("M0", "M1", "M2"), to = c("M0", "M1/2", "M1/2"))) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(STIM = case_when(STIM == "Wuhan E Wuhan M" ~ "Ancestral E/M",
                          STIM == "Wuhan N" ~ "Ancestral N",
                          STIM == "BA.1-2-4-5 E / BA.1-2-4-5 M" ~ "BA4/5 E/M",
                          STIM == "BA.4-5 N" ~ "BA4/5 N",
                          .default = STIM)) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- dt.ggplot.text (% of positive responses)
dt.ggplot.text <- dt.ggplot %>%
  mutate(MIMOSA_call = factor(x = MIMOSA_call, levels = c("FALSE", "TRUE"))) %>%
  group_by(RX_CODE, STIM, VISITMONTH, Leiden, MIMOSA_call, .drop = FALSE) %>%
  summarize(n = n()) %>%
  group_by(RX_CODE, STIM, VISITMONTH, Leiden) %>%
  mutate(N = sum(n)) %>%
  mutate(text = paste0(n, "/", N)) %>%
  mutate(text_2 = paste0(round(n/N*100, 0), "%")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  filter(!(str_detect(string = text, pattern = "NaN"))) %>%
  filter(text != "0/0") %>%
  mutate(COLOR_TXT = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1", "AG2-2") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- Fig. (1)
set.seed(1234)
cairo_pdf(filename = "boxplot-by-LEIDENxVISITMONTHxGROUP(1).pdf", 
    width = 6, 
    height = 6.25, 
    onefile = TRUE)
for(antigen in c("Any Ancestral", 
                 "Any BA4/5", 
                 "Spike Ancestral", 
                 "Spike BA4/5", 
                 "COV2 CON S1", 
                 "COV2 CON S2", 
                 "Ancestral N", 
                 "Ancestral E/M", 
                 "BA.4-5 S1", 
                 "BA.4-5 S2", 
                 "BA4/5 E/M", 
                 "BA4/5 N")){
  for(leiden in levels(dt$asinh_asym_Leiden)) {
    dt.ggplot.tmp <- dt.ggplot %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(X = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                           VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline_2",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination_2",
                           .default = "Post two\nvaccinations")) %>%
      mutate(X = factor(x = X, levels = c("Baseline", "Post two\nvaccinations", "Baseline_2", "Post one\nvaccination_2")))
    dt.ggplot.text.tmp <- dt.ggplot.text %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(X = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                           VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline_2",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination_2",
                           .default = "Post two\nvaccinations")) %>%
      mutate(X = factor(x = X, levels = c("Baseline", "Post two\nvaccinations", "Baseline_2", "Post one\nvaccination_2")))
    
    dt.ggplot.tmp %>%
      arrange(PTID) %>%
      ggplot() +
        geom_point(aes(x = X, y = PCTPOS_ADJ, fill = FILL_PCH, color = COLOR_PCH, group = PTID), 
                   position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                   size = 3, alpha = .7, pch = 21) +
        geom_line(aes(x = X, y = PCTPOS_ADJ, color = COLOR_LINE, group = PTID), 
                  position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                  linewidth = .2, alpha = .7) +
        geom_boxplot(aes(x = X, y = PCTPOS_ADJ, fill = FILL_BX),
                     alpha = .75, width = .5, outlier.shape = NA) +
        geom_text(aes(x = X, y = 15, label = text), 
                  size = 4,
                  data = dt.ggplot.text.tmp) +
        geom_text(aes(x = X, y = 8, label = text_2), 
                  color = "grey40", size = 3,
                  data = dt.ggplot.text.tmp) +
        geom_hline(aes(yintercept = 0),
                   color = "grey60", lty = 2, linewidth = .5,
                   data = dt.ggplot.tmp %>% filter(FACET_Y == "PLWH")) +
        facet_grid(FACET_Y ~ FACET_X, scales = "free_x", space = "free_x") +
        scale_x_discrete(breaks = c("Baseline", "Post two\nvaccinations", "Baseline_2", "Post one\nvaccination_2"),
                         labels = c("Baseline", "Post two\nvaccinations", "Baseline", "Post one\nvaccination"), 
                         drop = TRUE) +
        scale_y_continuous(trans = "log10", 
                           breaks = c(0.001, 0.01, 0.1, 1, 10), 
                           labels = c("\u22640.001", "0.01", "0.1", "1", "10"), 
                           limits = c(0.0008, 15)) +
        scale_fill_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                     "#B2DF8A" = "#B2DF8A", 
                                     "#FB9A99" = "#FB9A99", 
                                     "#FDBF6F" = "#FDBF6F", 
                                     "white" = "white")) +
        scale_color_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                      "#B2DF8A" = "#B2DF8A", 
                                      "#FB9A99" = "#FB9A99", 
                                      "#FDBF6F" = "#FDBF6F", 
                                      "white" = "white"), guide = "none") +
        labs(x = NULL, 
             y = "% of CD4+ T cells\nBackground-subtracted", 
             fill = NULL,
             title = paste(antigen, leiden, sep = "\n")) +
        theme_classic() +
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 14),
              strip.text = element_text(size = 16),
              title = element_text(size = 16),
              legend.position = "none") -> plot
    plot %>% print()
  }
}
dev.off()
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Box-plots
#- meta.data
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
   filter(FASFL == "Y") # 168 PTIDs
dt.Leiden <- dt.Leiden %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2N, 
                                   warn_missing = FALSE)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = RX_CODE, 
                                   from = c("1", "2.1", "3", "4.1"), 
                                   to = c("AG1", "AG2-1", "AG3", "AG4-1"), 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))  %>%
  mutate(COHORT = plyr::mapvalues(x = COHORT, 
                                  from = c("1", "0"), 
                                  to = c("case", "non-case"), 
                                  warn_missing = FALSE))

#- dt.ggplot
dt.ggplot <- dt.Leiden %>%
  ungroup() %>%
  filter(!(is.na(MIMOSA_call))) %>%
  mutate(VISITMONTH = factor(x = VISITMONTH, levels = c("M0", "M1", "M2"))) %>%
  mutate(FILL_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "TRUE" ~ "#A6CEE3",
                              RX_CODE %in% "AG2-1" & MIMOSA_call == "TRUE" ~ "#B2DF8A",
                              RX_CODE %in% "AG3" & MIMOSA_call == "TRUE" ~ "#FB9A99",
                              RX_CODE %in% "AG4-1" & MIMOSA_call == "TRUE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "FALSE" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" & MIMOSA_call == "FALSE" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" & MIMOSA_call == "FALSE" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" & MIMOSA_call == "FALSE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_LINE = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                                RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                                RX_CODE %in% "AG3" ~ "#FB9A99",
                                RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                                .default = "white")) %>%
  mutate(FILL_BX = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                             RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                             RX_CODE %in% "AG3" ~ "#FB9A99",
                             RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                             .default = "white")) %>%
  mutate(FILL_BX = factor(x = FILL_BX, levels = c("#FB9A99", "#FDBF6F", "#A6CEE3", "#B2DF8A"))) %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITMONTH, from = c("M0", "M1", "M2"), to = c("M0", "M1/2", "M1/2"))) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(STIM = case_when(STIM == "Wuhan E Wuhan M" ~ "Ancestral E/M",
                          STIM == "Wuhan N" ~ "Ancestral N",
                          STIM == "BA.1-2-4-5 E / BA.1-2-4-5 M" ~ "BA4/5 E/M",
                          STIM == "BA.4-5 N" ~ "BA4/5 N",
                          .default = STIM)) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- dt.ggplot.text (% of positive responses)
dt.ggplot.text <- dt.ggplot %>%
  mutate(MIMOSA_call = factor(x = MIMOSA_call, levels = c("FALSE", "TRUE"))) %>%
  group_by(RX_CODE, STIM, VISITMONTH, COHORT, Leiden, MIMOSA_call, .drop = FALSE) %>%
  summarize(n = n()) %>%
  group_by(RX_CODE, STIM, VISITMONTH, COHORT, Leiden) %>%
  mutate(N = sum(n)) %>%
  mutate(text = paste0(n, "/", N)) %>%
  mutate(text_2 = paste0(round(n/N*100, 0), "%")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  filter(!(str_detect(string = text, pattern = "NaN"))) %>%
  filter(text != "0/0") %>%
  mutate(COLOR_TXT = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1", "AG2-2") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- Fig. (2)
set.seed(1234)
cairo_pdf(filename = "boxplot-by-LEIDENxVISITMONTHxGROUP(2).pdf", 
    width = 10, 
    height = 7.25, 
    onefile = TRUE)
for(antigen in c("Any Ancestral", 
                 "Any BA4/5", 
                 "Spike Ancestral", 
                 "Spike BA4/5", 
                 "COV2 CON S1", 
                 "COV2 CON S2", 
                 "Ancestral N", 
                 "Ancestral E/M", 
                 "BA.4-5 S1", 
                 "BA.4-5 S2", 
                 "BA4/5 E/M", 
                 "BA4/5 N")){
  for(leiden in levels(dt$asinh_asym_Leiden)) {
    dt.ggplot.tmp <- dt.ggplot %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(FACET_X2 = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                                  VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination",
                                  .default = "Post two\nvaccinations")) %>%
      mutate(FACET_X2 = factor(x = FACET_X2, levels = c("Baseline", "Post two\nvaccinations", "Post one\nvaccination")))
    dt.ggplot.text.tmp <- dt.ggplot.text %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(X = case_when(COHORT == "case" ~ 1, COHORT == "non-case" ~ 2)) %>%
      mutate(FACET_X2 = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                                  VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination",
                                  .default = "Post two\nvaccinations")) %>%
      mutate(FACET_X2 = factor(x = FACET_X2, levels = c("Baseline", "Post two\nvaccinations", "Post one\nvaccination")))
      
    dt.ggplot.tmp %>%
      arrange(PTID) %>%
      ggplot() +
        geom_point(aes(x = COHORT, y = PCTPOS_ADJ, fill = FILL_PCH, color = COLOR_PCH, group = PTID), 
                   position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                   size = 3, alpha = .7, pch = 21) +
        geom_line(aes(x = COHORT, y = PCTPOS_ADJ, color = COLOR_LINE, group = PTID), 
                  position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                  linewidth = .2, alpha = .7) +
        geom_boxplot(aes(x = COHORT, y = PCTPOS_ADJ, fill = FILL_BX),
                     alpha = .75, width = .5, outlier.shape = NA) +
        geom_text(aes(x = X, y = 15, label = text), 
                  size = 4,
                  data = dt.ggplot.text.tmp) +
        geom_text(aes(x = X, y = 8, label = text_2), 
                  color = "grey40", size = 3,
                  data = dt.ggplot.text.tmp) +
        geom_hline(aes(yintercept = 0),
                   color = "grey60", lty = 2, linewidth = .5,
                   data = dt.ggplot.tmp %>% filter(FACET_Y == "PLWH")) +
        facet_grid(FACET_Y ~ FACET_X + FACET_X2, scales = "free_x", space = "free_x") +
        #scale_x_discrete(breaks = c("0", "1"), labels = c("non-case", "case")) +
        scale_y_continuous(trans = "log10", 
                           breaks = c(0.001, 0.01, 0.1, 1, 10), 
                           labels = c("\u22640.001", "0.01", "0.1", "1", "10"), 
                           limits = c(0.0009, 15)) +
        scale_fill_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                     "#B2DF8A" = "#B2DF8A", 
                                     "#FB9A99" = "#FB9A99", 
                                     "#FDBF6F" = "#FDBF6F", 
                                     "white" = "white")) +
        scale_color_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                      "#B2DF8A" = "#B2DF8A", 
                                      "#FB9A99" = "#FB9A99", 
                                      "#FDBF6F" = "#FDBF6F", 
                                      "white" = "white"), guide = "none") +
        labs(x = NULL, 
             y = "% of CD4+ T cells\nBackground-subtracted", 
             fill = NULL,
             title = paste(antigen, leiden, sep = "\n")) +
        theme_classic() +
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 14),
              strip.text = element_text(size = 16),
              title = element_text(size = 16),
              legend.position = "none") -> plot
    plot %>% print()
  }
}
dev.off()
```

**Primary Data Analysis**

Any cluster that has a response rate > 20% in at least one VISITNO in the vaccine group is kept for the Primary Data Analysis. This prioritization helps ensure to highlight vaccine-induced clusters.  

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Box-plots
#- meta.data
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
   filter(FASFL == "Y") # 168 PTIDs
dt.Leiden <- dt.Leiden %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2N, 
                                   warn_missing = FALSE)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = RX_CODE, 
                                   from = c("1", "2.1", "3", "4.1"), 
                                   to = c("AG1", "AG2-1", "AG3", "AG4-1"), 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))  %>%
  mutate(COHORT = plyr::mapvalues(x = COHORT, 
                                  from = c("1", "0"), 
                                  to = c("case", "non-case"), 
                                  warn_missing = FALSE))

#- Selection
dt.PDA <- dt.Leiden %>% 
  group_by(RX_CODE, STIM, VISITMONTH, Leiden, MIMOSA_call) %>%
  summarize(n = n()) %>%
  group_by(RX_CODE, STIM, VISITMONTH, Leiden) %>%
  mutate(N = sum(n)) %>% 
  mutate(freq = n / N) %>%
  filter(freq > .2 & MIMOSA_call == "TRUE")
leiden.clusters <- unique(dt.PDA$Leiden)
leiden.clusters %>% length() # 7 clusters (Leiden: 1, 2, 5, 6, 7, 9, 10)
leiden.stim <- paste(dt.PDA$Leiden, dt.PDA$STIM) %>% unique()
dt.PDA %>%
  filter(RX_CODE == "AG2-1") %>%
  View()

#- dt.ggplot
dt.ggplot <- dt.Leiden %>%
  ungroup() %>%
  filter(!(is.na(MIMOSA_call))) %>%
  filter(paste(Leiden, STIM) %in% leiden.stim) %>%
  mutate(Leiden = factor(x = Leiden, levels = paste("Leiden:", c(1, 2, 5, 6, 7, 9, 10)))) %>%
  mutate(VISITMONTH = factor(x = VISITMONTH, levels = c("M0", "M1", "M2"))) %>%
  mutate(FILL_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "TRUE" ~ "#A6CEE3",
                              RX_CODE %in% "AG2-1" & MIMOSA_call == "TRUE" ~ "#B2DF8A",
                              RX_CODE %in% "AG3" & MIMOSA_call == "TRUE" ~ "#FB9A99",
                              RX_CODE %in% "AG4-1" & MIMOSA_call == "TRUE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "FALSE" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" & MIMOSA_call == "FALSE" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" & MIMOSA_call == "FALSE" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" & MIMOSA_call == "FALSE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_LINE = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                                RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                                RX_CODE %in% "AG3" ~ "#FB9A99",
                                RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                                .default = "white")) %>%
  mutate(FILL_BX = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                             RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                             RX_CODE %in% "AG3" ~ "#FB9A99",
                             RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                             .default = "white")) %>%
  mutate(FILL_BX = factor(x = FILL_BX, levels = c("#FB9A99", "#FDBF6F", "#A6CEE3", "#B2DF8A"))) %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITMONTH, from = c("M0", "M1", "M2"), to = c("M0", "M1/2", "M1/2"))) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(STIM = case_when(STIM == "Wuhan E Wuhan M" ~ "Ancestral E/M",
                          STIM == "Wuhan N" ~ "Ancestral N",
                          STIM == "BA.1-2-4-5 E / BA.1-2-4-5 M" ~ "BA4/5 E/M",
                          STIM == "BA.4-5 N" ~ "BA4/5 N",
                          .default = STIM)) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- dt.ggplot.text (% of positive responses)
dt.ggplot.text <- dt.ggplot %>%
  mutate(MIMOSA_call = factor(x = MIMOSA_call, levels = c("FALSE", "TRUE"))) %>%
  group_by(RX_CODE, STIM, VISITMONTH, Leiden, MIMOSA_call, .drop = FALSE) %>%
  summarize(n = n()) %>%
  group_by(RX_CODE, STIM, VISITMONTH, Leiden) %>%
  mutate(N = sum(n)) %>%
  mutate(text = paste0(n, "/", N)) %>%
  mutate(text_2 = paste0(round(n/N*100, 0), "%")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  filter(!(str_detect(string = text, pattern = "NaN"))) %>%
  filter(text != "0/0") %>%
  mutate(COLOR_TXT = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1", "AG2-2") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- Fig. (1)
set.seed(1234)
cairo_pdf(filename = "boxplot-by-LEIDENxVISITMONTHxGROUP(1)_filt.pdf", 
    width = 6, 
    height = 6.25, 
    onefile = TRUE)
for(antigen in c("Any Ancestral", 
                 "Any BA4/5", 
                 "Spike Ancestral", 
                 "Spike BA4/5", 
                 "COV2 CON S1", 
                 "COV2 CON S2", 
                 "Ancestral N", 
                 "Ancestral E/M", 
                 "BA.4-5 S1", 
                 "BA.4-5 S2", 
                 "BA4/5 E/M", 
                 "BA4/5 N")){
  leiden.tmp <- dt.ggplot %>%
    filter(STIM == antigen) %>%
    pull(Leiden) %>%
    unique()
  for(leiden in leiden.tmp) {
    dt.ggplot.tmp <- dt.ggplot %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(X = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                           VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline_2",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination_2",
                           .default = "Post two\nvaccinations")) %>%
      mutate(X = factor(x = X, levels = c("Baseline", "Post two\nvaccinations", "Baseline_2", "Post one\nvaccination_2")))
    dt.ggplot.text.tmp <- dt.ggplot.text %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(X = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                           VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline_2",
                           VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination_2",
                           .default = "Post two\nvaccinations")) %>%
      mutate(X = factor(x = X, levels = c("Baseline", "Post two\nvaccinations", "Baseline_2", "Post one\nvaccination_2")))
    
    dt.ggplot.tmp %>%
      arrange(PTID) %>%
      ggplot() +
        geom_point(aes(x = X, y = PCTPOS_ADJ, fill = FILL_PCH, color = COLOR_PCH, group = PTID), 
                   position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                   size = 3, alpha = .7, pch = 21) +
        geom_line(aes(x = X, y = PCTPOS_ADJ, color = COLOR_LINE, group = PTID), 
                  position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                  linewidth = .2, alpha = .7) +
        geom_boxplot(aes(x = X, y = PCTPOS_ADJ, fill = FILL_BX),
                     alpha = .75, width = .5, outlier.shape = NA) +
        geom_text(aes(x = X, y = 15, label = text), 
                  size = 4,
                  data = dt.ggplot.text.tmp) +
        geom_text(aes(x = X, y = 8, label = text_2), 
                  color = "grey40", size = 3,
                  data = dt.ggplot.text.tmp) +
        geom_hline(aes(yintercept = 0),
                   color = "grey60", lty = 2, linewidth = .5,
                   data = dt.ggplot.tmp %>% filter(FACET_Y == "PLWH")) +
        facet_grid(FACET_Y ~ FACET_X, scales = "free_x", space = "free_x") +
        scale_x_discrete(breaks = c("Baseline", "Post two\nvaccinations", "Baseline_2", "Post one\nvaccination_2"),
                         labels = c("Baseline", "Post two\nvaccinations", "Baseline", "Post one\nvaccination"), 
                         drop = TRUE) +
        scale_y_continuous(trans = "log10", 
                           breaks = c(0.001, 0.01, 0.1, 1, 10), 
                           labels = c("\u22640.001", "0.01", "0.1", "1", "10"), 
                           limits = c(0.0009, 15)) +
        scale_fill_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                     "#B2DF8A" = "#B2DF8A", 
                                     "#FB9A99" = "#FB9A99", 
                                     "#FDBF6F" = "#FDBF6F", 
                                     "white" = "white")) +
        scale_color_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                      "#B2DF8A" = "#B2DF8A", 
                                      "#FB9A99" = "#FB9A99", 
                                      "#FDBF6F" = "#FDBF6F", 
                                      "white" = "white"), guide = "none") +
        labs(x = NULL, 
             y = "% of CD4+ T cells\nBackground-subtracted", 
             fill = NULL,
             title = paste(antigen, leiden, sep = "\n")) +
        theme_classic() +
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 14),
              strip.text = element_text(size = 16),
              title = element_text(size = 16),
              legend.position = "none") -> plot
    plot %>% print()
  }
}
dev.off()
```
```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
###--- Box-plots
#- meta.data
meta.data <- read_csv("./COVID_realdata_ics_bama_nab_bcp20241023_processed_with_riskscore.csv") %>%
  filter(ph2.m1.ics.cohort == "1") %>%
   filter(FASFL == "Y") # 168 PTIDs
dt.Leiden <- dt.Leiden %>%
  mutate(RX_CODE = plyr::mapvalues(x = PTID, 
                                   from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                   to = meta.data$AGRP2N, 
                                   warn_missing = FALSE)) %>%
  mutate(RX_CODE = plyr::mapvalues(x = RX_CODE, 
                                   from = c("1", "2.1", "3", "4.1"), 
                                   to = c("AG1", "AG2-1", "AG3", "AG4-1"), 
                                   warn_missing = FALSE)) %>%
  mutate(COHORT = plyr::mapvalues(x = PTID, 
                                  from = str_remove(string = meta.data$Subjectid, pattern = "C3008[-]"),
                                  to = meta.data$EventIndPrimary1, 
                                  warn_missing = FALSE))  %>%
  mutate(COHORT = plyr::mapvalues(x = COHORT, 
                                  from = c("1", "0"), 
                                  to = c("case", "non-case"), 
                                  warn_missing = FALSE))

#- Selection
dt.PDA <- dt.Leiden %>% 
  group_by(RX_CODE, STIM, VISITMONTH, Leiden, MIMOSA_call) %>%
  summarize(n = n()) %>%
  group_by(RX_CODE, STIM, VISITMONTH, Leiden) %>%
  mutate(N = sum(n)) %>% 
  mutate(freq = n / N) %>%
  filter(freq > .2 & MIMOSA_call == "TRUE")
leiden.clusters <- unique(dt.PDA$Leiden)
leiden.clusters %>% length() # 7 clusters (Leiden: 1, 2, 5, 6, 7, 9, 10)
leiden.stim <- paste(dt.PDA$Leiden, dt.PDA$STIM) %>% unique()

#- dt.ggplot
dt.ggplot <- dt.Leiden %>%
  ungroup() %>%
  filter(!(is.na(MIMOSA_call))) %>%
  filter(paste(Leiden, STIM) %in% leiden.stim) %>%
  mutate(Leiden = factor(x = Leiden, levels = paste("Leiden:", c(1, 2, 5, 6, 7, 9, 10)))) %>%
  mutate(VISITMONTH = factor(x = VISITMONTH, levels = c("M0", "M1", "M2"))) %>%
  mutate(FILL_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "TRUE" ~ "#A6CEE3",
                              RX_CODE %in% "AG2-1" & MIMOSA_call == "TRUE" ~ "#B2DF8A",
                              RX_CODE %in% "AG3" & MIMOSA_call == "TRUE" ~ "#FB9A99",
                              RX_CODE %in% "AG4-1" & MIMOSA_call == "TRUE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_PCH = case_when(RX_CODE %in% "AG1" & MIMOSA_call == "FALSE" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" & MIMOSA_call == "FALSE" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" & MIMOSA_call == "FALSE" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" & MIMOSA_call == "FALSE" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(COLOR_LINE = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                                RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                                RX_CODE %in% "AG3" ~ "#FB9A99",
                                RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                                .default = "white")) %>%
  mutate(FILL_BX = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                             RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                             RX_CODE %in% "AG3" ~ "#FB9A99",
                             RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                             .default = "white")) %>%
  mutate(FILL_BX = factor(x = FILL_BX, levels = c("#FB9A99", "#FDBF6F", "#A6CEE3", "#B2DF8A"))) %>%
  mutate(VISITMONTH = plyr::mapvalues(x = VISITMONTH, from = c("M0", "M1", "M2"), to = c("M0", "M1/2", "M1/2"))) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(STIM = case_when(STIM == "Wuhan E Wuhan M" ~ "Ancestral E/M",
                          STIM == "Wuhan N" ~ "Ancestral N",
                          STIM == "BA.1-2-4-5 E / BA.1-2-4-5 M" ~ "BA4/5 E/M",
                          STIM == "BA.4-5 N" ~ "BA4/5 N",
                          .default = STIM)) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- dt.ggplot.text (% of positive responses)
dt.ggplot.text <- dt.ggplot %>%
  mutate(MIMOSA_call = factor(x = MIMOSA_call, levels = c("FALSE", "TRUE"))) %>%
  group_by(RX_CODE, STIM, VISITMONTH, COHORT, Leiden, MIMOSA_call, .drop = FALSE) %>%
  summarize(n = n()) %>%
  group_by(RX_CODE, STIM, VISITMONTH, COHORT, Leiden) %>%
  mutate(N = sum(n)) %>%
  mutate(text = paste0(n, "/", N)) %>%
  mutate(text_2 = paste0(round(n/N*100, 0), "%")) %>%
  filter(MIMOSA_call == "TRUE") %>%
  filter(!(str_detect(string = text, pattern = "NaN"))) %>%
  filter(text != "0/0") %>%
  mutate(COLOR_TXT = case_when(RX_CODE %in% "AG1" ~ "#A6CEE3",
                               RX_CODE %in% "AG2-1" ~ "#B2DF8A",
                               RX_CODE %in% "AG3" ~ "#FB9A99",
                               RX_CODE %in% "AG4-1" ~ "#FDBF6F",
                              .default = "white")) %>%
  mutate(RX_CODE = factor(x = RX_CODE, levels = c("AG3", "AG4-1", "AG1", "AG2-1"))) %>%
  mutate(FACET_X = case_when(RX_CODE %in% c("AG1", "AG3") ~ "SARS2-", .default = "SARS2+")) %>%
  mutate(FACET_X = factor(x = FACET_X, levels = c("SARS2-", "SARS2+"))) %>%
  mutate(FACET_Y = case_when(RX_CODE %in% c("AG1", "AG2-1", "AG2-2") ~ "PLWH", .default = "PWoH")) %>%
  mutate(FACET_Y = factor(x = FACET_Y, levels = c("PLWH", "PWoH")))

#- Fig. (2)
set.seed(1234)
cairo_pdf(filename = "boxplot-by-LEIDENxVISITMONTHxGROUP(2)_filt.pdf", 
    width = 10, 
    height = 7.25, 
    onefile = TRUE)
for(antigen in c("Any Ancestral", 
                 "Any BA4/5", 
                 "Spike Ancestral", 
                 "Spike BA4/5", 
                 "COV2 CON S1", 
                 "COV2 CON S2", 
                 "Ancestral N", 
                 "Ancestral E/M", 
                 "BA.4-5 S1", 
                 "BA.4-5 S2", 
                 "BA4/5 E/M", 
                 "BA4/5 N")){
  leiden.tmp <- dt.ggplot %>%
    filter(STIM == antigen) %>%
    pull(Leiden) %>%
    unique()
  for(leiden in leiden.tmp) {
    dt.ggplot.tmp <- dt.ggplot %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(FACET_X2 = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                                  VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination",
                                  .default = "Post two\nvaccinations")) %>%
      mutate(FACET_X2 = factor(x = FACET_X2, levels = c("Baseline", "Post two\nvaccinations", "Post one\nvaccination")))
    dt.ggplot.text.tmp <- dt.ggplot.text %>%
      filter(STIM == antigen) %>%
      filter(Leiden == leiden) %>%
      mutate(X = case_when(COHORT == "case" ~ 1, COHORT == "non-case" ~ 2)) %>%
      mutate(FACET_X2 = case_when(VISITMONTH == "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG1", "AG3") ~ "Post two\nvaccinations",
                                  VISITMONTH == "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Baseline",
                                  VISITMONTH != "M0" & RX_CODE %in% c("AG4-1", "AG2-1") ~ "Post one\nvaccination",
                                  .default = "Post two\nvaccinations")) %>%
      mutate(FACET_X2 = factor(x = FACET_X2, levels = c("Baseline", "Post two\nvaccinations", "Post one\nvaccination")))
      
    dt.ggplot.tmp %>%
      arrange(PTID) %>%
      ggplot() +
        geom_point(aes(x = COHORT, y = PCTPOS_ADJ, fill = FILL_PCH, color = COLOR_PCH, group = PTID), 
                   position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                   size = 3, alpha = .7, pch = 21) +
        geom_line(aes(x = COHORT, y = PCTPOS_ADJ, color = COLOR_LINE, group = PTID), 
                  position = position_jitterdodge(jitter.width = .05, dodge.width = .4, seed = 123),
                  linewidth = .2, alpha = .7) +
        geom_boxplot(aes(x = COHORT, y = PCTPOS_ADJ, fill = FILL_BX),
                     alpha = .75, width = .5, outlier.shape = NA) +
        geom_text(aes(x = X, y = 15, label = text), 
                  size = 4,
                  data = dt.ggplot.text.tmp) +
        geom_text(aes(x = X, y = 8, label = text_2), 
                  color = "grey40", size = 3,
                  data = dt.ggplot.text.tmp) +
        geom_hline(aes(yintercept = 0),
                   color = "grey60", lty = 2, linewidth = .5,
                   data = dt.ggplot.tmp %>% filter(FACET_Y == "PLWH")) +
        facet_grid(FACET_Y ~ FACET_X + FACET_X2, scales = "free_x", space = "free_x") +
        #scale_x_discrete(breaks = c("0", "1"), labels = c("non-case", "case")) +
        scale_y_continuous(trans = "log10", 
                           breaks = c(0.001, 0.01, 0.1, 1, 10), 
                           labels = c("\u22640.001", "0.01", "0.1", "1", "10"), 
                           limits = c(0.0009, 15)) +
        scale_fill_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                     "#B2DF8A" = "#B2DF8A", 
                                     "#FB9A99" = "#FB9A99", 
                                     "#FDBF6F" = "#FDBF6F", 
                                     "white" = "white")) +
        scale_color_manual(values = c("#A6CEE3" = "#A6CEE3", 
                                      "#B2DF8A" = "#B2DF8A", 
                                      "#FB9A99" = "#FB9A99", 
                                      "#FDBF6F" = "#FDBF6F", 
                                      "white" = "white"), guide = "none") +
        labs(x = NULL, 
             y = "% of CD4+ T cells\nBackground-subtracted", 
             fill = NULL,
             title = paste(antigen, leiden, sep = "\n")) +
        theme_classic() +
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 14),
              strip.text = element_text(size = 16),
              title = element_text(size = 16),
              legend.position = "none") -> plot
    plot %>% print()
  }
}
dev.off()
```

