---
title: "Coxph Report Serum & ICS"
author: "CoVPN 3008"
date: "2024-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
# Change the directory as needed
setwd('Github/Tcell_manuscript/Cox_HR_ICS/')
knitr::opts_chunk$set(echo = TRUE)
set.seed(1)
```


```{r load pckages, echo=FALSE, warning=FALSE, message=FALSE}
library(CaseCohortCoxSurvival)
library(survival)
library(dplyr)
library(survey)
library(ggplot2)
library(vaccine)
library(wCorr)
library(coxphw)
library(haven)
library(GGally)
library(Hmisc)
source('coxph_functions_PBMC.R')
```

# Data preprocessing

```{r data wrangling, echo=FALSE}
dt = read.csv('./COVID_realdata_ics_bama_nab_bcp20241210_processed_with_riskscore.csv')

# PBMC correlates analysis is restricted to PLWH
dt_PBMC_PP = dt %>% 
  filter(ph1.m1.ics == 1) %>%
  mutate(hybrid = (AGRP2N == 2.1) + 0) %>%
  mutate(is.subcohort = ifelse(ph2.m1.ics == 1 & IMMFL == 'Y', 1, 0)) %>%
   mutate(log10M1BA45ID50 = log10(M1BA45ID50),
         log10BBA45ID50 = log10(BBA45ID50),
         frBA45ID50 = log10M1BA45ID50 - log10BBA45ID50,
         log10M1IGGSpikeIndex = log10(M1IGGMSDA16AU), #A16 = IgG S Index
         log10BIGGSpikeIndex = log10(BIGGMSDA16AU), 
         frIGGSpikeIndex = log10M1IGGSpikeIndex - log10BIGGSpikeIndex,
         log10M1IGGSpikeBA45 = log10(M1IGGMSDA23AU), # A23 = IgG S BA.4/5
         log10BIGGSpikeBA45 = log10(BIGGMSDA23AU), 
         frIGGSpikeBA45 = log10M1IGGSpikeBA45 - log10BIGGSpikeBA45,
         log10M1IGGN = log10(M1IGGMSDA15AU),
         log10BIGGN = log10(BIGGMSDA15AU))

# Define a stratification variable W corresponding to
# 4 baseline strata + case
# Mutate ph2 to logical 
dt_PBMC_PP = dt_PBMC_PP %>%
  mutate(W = ifelse(EventIndPrimary1 == 1, 1, 
                    ifelse(AGRP2N == 1, 2, 3))) %>%
  mutate(ph2.m1.ics_logical = ifelse(ph2.m1.ics == 1, TRUE, FALSE))

dt_PBMC_PP = dt_PBMC_PP %>%
  mutate(TBFL = ifelse(TBFL == '', 'N', TBFL),
         TBFL = ifelse(TBFL == 'Y', 1, 0)) %>%
  mutate(AGEGR1 = ifelse(AGEGR1 == '>40', 1, 0),
         BMIB2GR1 = ifelse(BMIB2GR1 == '>25', 1, 0),
         CD4B2GR1 = ifelse(is.na(CD4B2GR1), 'NotKnown', 
                           ifelse(CD4B2GR1 == '<=500', 'low', 'high')),
         riskscore = ifelse(is.na(riskscore), median(riskscore, na.rm = T), riskscore),
         enrollmentPeriodEarly = ifelse(CalendarDateEnrollment <= 90, 1, 0),
         enrollmentPeriodMid = ifelse((CalendarDateEnrollment > 90 & CalendarDateEnrollment <= 180), 1, 0),
         regionRSA = ifelse(REGION == 'South Africa', 1, 0))
```


## Define serum markers

```{r setmarkerofinterest serum, echo=TRUE}
antigen_vec_serum = c('BA45ID50', 'IGGSpikeIndex', 'IGGSpikeBA45', 'IGGN')
antigen_names_vec_serum = c('nAb ID50 BA.4/5', 'IgG S Index',  
                      'IgG S BA.4/5', 'IgG N Index')
mk_vec_b_serum = paste0('log10B', antigen_vec_serum)
names_vec_b_serum = c('M0 nAb-ID50 BA.4/5 (AU/ml)',
              'M0 IgG S Index (AU/ml)',
              'M0 IgG S BA.4/5 (AU/ml)',
              'M0 IgG N Index (AU/ml)')

mk_vec_p_serum = paste0('log10M1', antigen_vec_serum)
names_vec_p_serum = c('M1 nAb-ID50 BA.4/5 (AU/ml)',
              'M1 IgG S Index (AU/ml)',
              'M1 IgG S BA.4/5 (AU/ml)',
              'M1 IgG N Index (AU/ml)')

mk_vec_fr_serum = c('frBA45ID50', 'frIGGSpikeIndex', 'frIGGSpikeBA45')
names_vec_fr_serum = c('Fold-rise nAb-ID50 BA.4/5 (AU/ml)',
              'Fold-rise IgG S Index (AU/ml)',
              'Fold-rise IgG S BA.4/5 (AU/ml)')
```

## Define ICS markers

```{r setmarkerofinterest T, echo=TRUE}
# A11 = Spike BA.4/5; A02 = BA.4/5 N; A10 = Any BA.4/5
# C03: 'IFNg OR IL-2'; C04: 'IFNg OR IL-2 OR CD154'
T_vec = c('T4')
T_names_vec = c('CD4+')
C_vec = c('C03', 'C04')
C_names_vec = c('IFNg OR IL-2', 'IFNg OR IL-2 OR CD154')
A_vec = c('A02', 'A11')
A_names_vec = c('N BA.4/5', 'S BA.4/5')

antigen_vec_T = NULL
antigen_names_vec_T = NULL
for (i in 1:length(T_vec)) {
  for (j in 1:length(C_vec)) {
    for (k in 1:length(A_vec)) {
      antigen_vec_T = c(antigen_vec_T,
                      paste0(T_vec[i], C_vec[j], A_vec[k]))
      antigen_names_vec_T = c(antigen_names_vec_T,
                            paste0(T_names_vec[i], ' ', C_names_vec[j], ' ', A_names_vec[k]))
    }
  }
} 

mk_vec_b_T = paste0('B', antigen_vec_T)
names_vec_b_T = paste0('M0 ', antigen_names_vec_T, ' (%)')

mk_vec_p_T = paste0('M1', antigen_vec_T)
names_vec_p_T = paste0('M1 ', antigen_names_vec_T, ' (%)')

process <- function(x) log10(ifelse(x < 0.01, 0.01, x))

dt_PBMC_PP = dt_PBMC_PP %>%
  mutate_at(mk_vec_b_T, process) %>%
  mutate_at(mk_vec_p_T, process)

# Add fold-rise markers
for (antigen in antigen_vec_T) {
  dt_PBMC_PP = dt_PBMC_PP %>%
    mutate(!!paste0('fr', antigen) :=  !!sym(paste0('M1', antigen)) - !!sym(paste0('B', antigen)))
}
```

```{r standardize marker, echo = FALSE}
dt_PBMC_PP = dt_PBMC_PP %>%
  mutate(BT4C03A02 = ifelse(ph2.m1.ics == 1, BT4C03A02, NA),
         M1T4C03A02 = ifelse(ph2.m1.ics == 1, M1T4C03A02, NA),
         BT4C04A02 = ifelse(ph2.m1.ics == 1, BT4C04A02, NA),
         M1T4C04A02 = ifelse(ph2.m1.ics == 1, M1T4C04A02, NA),
         BT4C03A11 = ifelse(ph2.m1.ics == 1, BT4C03A11, NA),
         M1T4C03A11 = ifelse(ph2.m1.ics == 1, M1T4C03A11, NA),
         BT4C04A11 = ifelse(ph2.m1.ics == 1, BT4C04A11, NA),
         M1T4C04A11 = ifelse(ph2.m1.ics == 1, M1T4C04A11, NA))

# Standardized each ICS marker
dt_PBMC_PP = dt_PBMC_PP %>%
  mutate(BT4C03A02 = (BT4C03A02 - wtd.mean(BT4C03A02, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(BT4C03A02, weights = wt.m1.ics, na.rm =TRUE)),
         M1T4C03A02 = (M1T4C03A02 - wtd.mean(M1T4C03A02, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(M1T4C03A02, weights = wt.m1.ics, na.rm =TRUE)),
      
         
         BT4C03A11 = (BT4C03A11 - wtd.mean(BT4C03A11, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(BT4C03A11, weights = wt.m1.ics, na.rm =TRUE)),
         M1T4C03A11 = (M1T4C03A11 - wtd.mean(M1T4C03A11, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(M1T4C03A11, weights = wt.m1.ics, na.rm =TRUE)),
        
         BT4C04A02 = (BT4C04A02 - wtd.mean(BT4C04A02, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(BT4C04A02, weights = wt.m1.ics, na.rm =TRUE)),
         M1T4C04A02 = (M1T4C04A02 - wtd.mean(M1T4C04A02, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(M1T4C04A02, weights = wt.m1.ics, na.rm =TRUE)),
        
         
         BT4C04A11 = (BT4C04A11 - wtd.mean(BT4C04A11, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(BT4C04A11, weights = wt.m1.ics, na.rm =TRUE)),
         M1T4C04A11 = (M1T4C04A11 - wtd.mean(M1T4C04A11, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(M1T4C04A11, weights = wt.m1.ics, na.rm =TRUE)))


dt_PBMC_PP = dt_PBMC_PP %>%
  mutate(log10M1BA45ID50 = ifelse(ph2.m1.ics == 1, log10M1BA45ID50, NA),
         log10BBA45ID50 = ifelse(ph2.m1.ics == 1, log10BBA45ID50, NA),
         frBA45ID50 = ifelse(ph2.m1.ics == 1, frBA45ID50, NA),
         log10M1IGGSpikeIndex = ifelse(ph2.m1.ics == 1, log10M1IGGSpikeIndex, NA),
         log10BIGGSpikeIndex = ifelse(ph2.m1.ics == 1, log10BIGGSpikeIndex, NA),
         frIGGSpikeIndex = ifelse(ph2.m1.ics == 1, frIGGSpikeIndex, NA),
         log10M1IGGSpikeBA45 = ifelse(ph2.m1.ics == 1, log10M1IGGSpikeBA45, NA),
         log10BIGGSpikeBA45 = ifelse(ph2.m1.ics == 1, log10BIGGSpikeBA45, NA),
         frIGGSpikeBA45 = ifelse(ph2.m1.ics == 1, frIGGSpikeBA45, NA),
         log10M1IGGN = ifelse(ph2.m1.ics == 1, log10M1IGGN, NA),
         log10BIGGN = ifelse(ph2.m1.ics == 1, log10BIGGN, NA))

  # We further standard each M0, peak and fold-rise serum marker
dt_PBMC_PP = dt_PBMC_PP %>%
  mutate(log10M1BA45ID50 = (log10M1BA45ID50 - wtd.mean(log10M1BA45ID50, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10M1BA45ID50, weights = wt.m1.ics, na.rm =TRUE)),
         log10BBA45ID50 = (log10BBA45ID50 - wtd.mean(log10BBA45ID50, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10BBA45ID50, weights = wt.m1.ics, na.rm =TRUE)),
         frBA45ID50 = (frBA45ID50 - wtd.mean(frBA45ID50, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(frBA45ID50, weights = wt.m1.ics, na.rm =TRUE)),
         log10M1IGGSpikeIndex = (log10M1IGGSpikeIndex - wtd.mean(log10M1IGGSpikeIndex, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10M1IGGSpikeIndex, weights = wt.m1.ics, na.rm =TRUE)),
         log10BIGGSpikeIndex = (log10BIGGSpikeIndex - wtd.mean(log10BIGGSpikeIndex, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10BIGGSpikeIndex, weights = wt.m1.ics, na.rm =TRUE)),
         frIGGSpikeIndex = (frIGGSpikeIndex - wtd.mean(frIGGSpikeIndex, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(frIGGSpikeIndex, weights = wt.m1.ics, na.rm =TRUE)),
          log10M1IGGSpikeBA45 = (log10M1IGGSpikeBA45 - wtd.mean(log10M1IGGSpikeBA45, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10M1IGGSpikeBA45, weights = wt.m1.ics, na.rm =TRUE)),
          log10BIGGSpikeBA45 = (log10BIGGSpikeBA45 - wtd.mean(log10BIGGSpikeBA45, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10BIGGSpikeBA45, weights = wt.m1.ics, na.rm =TRUE)),
          frIGGSpikeBA45 = (frIGGSpikeBA45 - wtd.mean(frIGGSpikeBA45, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(frIGGSpikeBA45, weights = wt.m1.ics, na.rm =TRUE)),
         log10M1IGGN = (log10M1IGGN - wtd.mean(log10M1IGGN, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10M1IGGN, weights = wt.m1.ics, na.rm =TRUE)),
          log10BIGGN = (log10BIGGN - wtd.mean(log10BIGGN, weights = wt.m1.ics, na.rm =TRUE))/sqrt(wtd.var(log10BIGGN, weights = wt.m1.ics, na.rm =TRUE)))
```

## Define adjusted covariates
```{r adj covs}
adj_vars = c('riskscore') 
n_var = 1
```

# Hybrid immunity group correlates of risk analysis

```{r cohort HIPP}
dt_PBMC_PP_HIPP = dt_PBMC_PP %>%
  filter(hybrid == 1)
```

## Bivariate Correlates of Risk Analysis: adjusting for one baseline serum marker and one M0 ICS marker

```{r bivariate 1 HIPP, echo = TRUE, warning=FALSE, message=FALSE}
for (i in 1:length(mk_vec_b_serum)){
  for (j in 1:length(mk_vec_b_T)){
  md <- case_cohort_Cox_1(data = dt_PBMC_PP_HIPP, caseInd = "EventIndPrimary1", 
                            subcohortInd = "is.subcohort", 
                            baseline_strata  = "AGRP2N", 
                            time_to_event = 'EventTimePrimary', 
                            phase1_var_list = adj_vars, 
                            phase2_var_list  = c(mk_vec_b_serum[i], mk_vec_b_T[j]))
  coxfit = process_cccs_output(md)
  res = coxfit[(nrow(coxfit) - 1):nrow(coxfit),]
  res$term = c(names_vec_b_serum[i], names_vec_b_T[j])
  print(knitr::kable(res, digits = 3, row.names = FALSE, 
             col.names = c('Covariate', 'HR', 'P-Value', 'Lower CI', 'Upper CI')))
  wald_test_p = coxphw::wald(coeff = md$coxph.fit$coefficients, 
                    cov = md$coxph.fit$var, 
                    index = c((nrow(coxfit) - 1):nrow(coxfit)))['p']
  cat('Generalized Wald Test of ', names_vec_b_serum[i], ' and ', names_vec_b_T[j],'\n',  
  'P = ', round(wald_test_p, digits = 3) , '\n')
  }
}

```

## Bivariate Correlates of Risk Analysis: adjusting for one baseline serum marker and one peak ICS marker

```{r bivariate 2 HIPP, echo = TRUE, warning=FALSE, message=FALSE}
for (i in 1:length(mk_vec_b_serum)){
  for (j in 1:length(mk_vec_p_T)){
  md <- case_cohort_Cox_1(data = dt_PBMC_PP_HIPP, caseInd = "EventIndPrimary1", 
                            subcohortInd = "is.subcohort", 
                            baseline_strata  = "AGRP2N", 
                            time_to_event = 'EventTimePrimary', 
                            phase1_var_list = adj_vars, 
                            phase2_var_list  = c(mk_vec_b_serum[i], mk_vec_p_T[j]))
  coxfit = process_cccs_output(md)
  res = coxfit[(nrow(coxfit) - 1):nrow(coxfit),]
  res$term = c(names_vec_b_serum[i], names_vec_p_T[j])
  print(knitr::kable(res, digits = 3, row.names = FALSE, 
             col.names = c('Covariate', 'HR', 'P-Value', 'Lower CI', 'Upper CI')))
  wald_test_p = coxphw::wald(coeff = md$coxph.fit$coefficients, 
                    cov = md$coxph.fit$var, 
                    index = c((nrow(coxfit) - 1):nrow(coxfit)))['p']
  cat('Generalized Wald Test of ', names_vec_b_serum[i], ' and ', names_vec_p_T[j],'\n',  
  'P = ', round(wald_test_p, digits = 3) , '\n')
  }
}

```

## Partial likelihood ratio test

```{r analysis of deviance, echo = FALSE}
data = dt_PBMC_PP_HIPP
CaseCohortDesign<-twophase(id=list(~Subjectid,~Subjectid), 
                             strata=list(NULL,~W),
                             subset=~ph2.m1.ics_logical, data = data)
  
 # Prepare the formula
 mk = c('log10BBA45ID50', 'M1T4C04A11')
 RHS = paste(c(adj_vars, mk), collapse = '+')
 f = as.formula(paste0('Surv(EventTimePrimary, EventIndPrimary1) ~', RHS))
 
 # md is a model with baseline risk score, baseline nAb ID-50, and Peak ICS expressing
 # IFNg OR IL-2 OR CD154 against Spike BA.4/5
 md = survival::coxph(formula = f, weights = 1/CaseCohortDesign$prob, data =
                         data[data$ph2.m1.ics == 1,])
 
 # Test if the coef of M0 nAb ID-50 is 0 in md
 regTermTest(md, test.terms = 'log10BBA45ID50', method = 'LRT')
 
 # Test if the coef of Peak T4C04A11 is 0 in md
 regTermTest(md, test.terms = 'M1T4C04A11', method = 'LRT')
 
 
 # Prepare the formula
 mk = c('log10BBA45ID50', 'M1T4C03A11')
 RHS = paste(c(adj_vars, mk), collapse = '+')
 f = as.formula(paste0('Surv(EventTimePrimary, EventIndPrimary1) ~', RHS))

 # md is a model with baseline risk score, baseline nAb ID-50, and Peak ICS expressing
 # IFNg OR IL-2 against Spike BA.4/5
 md2 = survival::coxph(formula = f, weights = 1/CaseCohortDesign$prob, data =
                         data[data$ph2.m1.ics == 1,])
 
 # Test if the coef of M0 nAb ID-50 is 0 in md2
 regTermTest(md2, test.terms = 'log10BBA45ID50', method = 'LRT')
 
 # Test if the coef of Peak T4C03A11 is 0 in md2
 regTermTest(md2, test.terms = 'M1T4C03A11', method = 'LRT')
 
```
